---
title: "Combining_datasets_STdeconvolve_SR"
author: "Sofia Rouot"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: html_document
format:
  html:
    code-fold: true
    toc: true
    number-sections: true
    theme: spacelab
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, warning = FALSE}
library(STdeconvolve)
library(Seurat)
library(SeuratObject)
library(openxlsx)
library(RColorBrewer)
library(ggplot2)
```


# Spruce VegOct data

Load data
```{r}
se_spruce <- readRDS(file = "spruce_compatible_obj.all_samples.rds")
```

```{r}
se_spruce@meta.data
se_spruce.unique.samples <- unique(se_spruce$section_id)
se_spruce.unique.samples 
length(se_spruce.unique.samples) #88

se_spruce.unique.stages <- unique(se_spruce$bud)
se_spruce.unique.stages
length(se_spruce.unique.stages) #29 unique bud stages
```

Split for the Veg stage
```{r}
se_spruce_Veg <- subset(x = se_spruce, subset = stage == "Veg")
```

```{r}
length(unique(se_spruce_Veg$section_id))
```
37 section ids of Veg 

```{r}
length(unique(se_spruce_Veg$bud))
q <-unique(se_spruce_Veg$bud)
q
```

11 Veg buds (OctVeg1 etc)


Split for the October timepoint
```{r}
se_spruce_VegOct <- subset(x = se_spruce_Veg, subset = timepoint == "Oct")
se_spruce_VegOct@meta.data

unique(se_spruce_VegOct$timepoint)
```

Biological replicates
```{r}
unique(se_spruce_VegOct$bud)
```
3 biological replicates: "OctVeg3" "OctVeg1" "OctVeg2"

Technical replicates 
```{r}
unique(se_spruce_VegOct$section_id)
```
9 section ids : "V10F24-117-C1-1" "V10F24-117-C1-2" "V10F24-117-C1-3" "V19T26-078-A1-1" "V19T26-078-A1-2" "V19T26-078-A1-3"
"V19T26-078-B1-1" "V19T26-078-B1-2" "V19T26-078-B1-3"


Split into a list of section ids 
```{r}
se_spruce_VegOct_list <- SplitObject(se_spruce_VegOct, split.by = "section_id")
se_spruce_VegOct_list[[1]]@meta.data
se_spruce_VegOct_list[[9]]@meta.data
```

## Create corpus for each section id subsetted seurat object
N.B. The counts matrix are originally:
  - rows = genes 
  - columns = pixels 
  But STdeconvolve assumes the other way:
  - rows = pixels 
  - columns = genes
  
--> Use as.matrix(t(counts)) to transpose the matrix and ensure it is treated as a matrix. 
```{r}
VegOctCorpusReps_v2 <- lapply(se_spruce_VegOct_list, function(seurat_obj) {
  counts <- seurat_obj@assays[["Spatial"]]@counts
  #with transpose of matrix 
  dat <- preprocess(as.matrix(t(counts)),
                   extractPos = FALSE,
                   selected.genes = NA,
                   nTopGenes = NA,
                   genes.to.remove = NA,
                   removeAbove = 1.1,
                   removeBelow = 0.01,
                   min.reads = 1,
                   min.lib.size = 50,
                   min.detected = 1,
                   ODgenes = TRUE,
                   nTopOD = 1000,
                   od.genes.alpha = 0.05,
                   gam.k = 5) 
  dat
})
```

Add the x,y positions manually in the corpus 
```{r}
unique_section_id <- unique(se_spruce_VegOct$section_id)
for (section_id in unique_section_id){
  # Access your Seurat object
  seurat_obj <- se_spruce_VegOct_list[[section_id]]
  
  # Step 1: Get the counts matrix and its pixel names
  counts <- seurat_obj@assays$Spatial@counts
  
  # Step 2: Get the true pixel coordinates (a matrix or data.frame of x/y positions)
  image_name <- names(se_spruce_VegOct_list[[section_id]]@images)
  coords <- seurat_obj@images[[image_name]]@boundaries[["centroids"]]@coords
  
  # Make sure coords rownames match column names of counts
  rownames(coords) <- colnames(counts)
  
  # Step 4: Assign coords to pos slot of corpus
  VegOctCorpusReps_v2[[section_id]]$pos <- coords

}

```

Replace "-" with "_"
```{r}
names(VegOctCorpusReps_v2) <- c("V10F24_117_C1_1", "V10F24_117_C1_2", "V10F24_117_C1_3", "V19T26_078_A1_1", "V19T26_078_A1_2", "V19T26_078_A1_3", "V19T26_078_B1_1", "V19T26_078_B1_2", "V19T26_078_B1_3")
```


```{r}
combinedVegOctgenes <- Reduce(union, list(colnames(VegOctCorpusReps_v2$V10F24_117_C1_1$corpus),
                                       colnames(VegOctCorpusReps_v2$V10F24_117_C1_2$corpus),
                                       colnames(VegOctCorpusReps_v2$V10F24_117_C1_3$corpus),
                                       colnames(VegOctCorpusReps_v2$V19T26_078_A1_1$corpus),
                                       colnames(VegOctCorpusReps_v2$V19T26_078_A1_2$corpus),
                                       colnames(VegOctCorpusReps_v2$V19T26_078_A1_3$corpus),
                                       colnames(VegOctCorpusReps_v2$V19T26_078_B1_1$corpus),
                                       colnames(VegOctCorpusReps_v2$V19T26_078_B1_2$corpus),
                                       colnames(VegOctCorpusReps_v2$V19T26_078_B1_3$corpus))
                           )
length(combinedVegOctgenes)
```
3340 combined genes used for STdeconvolve of VegOct samples

```{r}
unique_section_id
```

Create the combined corpus with these 3340 genes and the (x.y) positions of each section_id 
```{r}
combinedVegOctcorpus <- dplyr::bind_rows(
  as.data.frame(t(se_spruce_VegOct_list[["V10F24-117-C1-1"]]@assays[["Spatial"]]@counts))[rownames(VegOctCorpusReps_v2$V10F24_117_C1_1$pos),combinedVegOctgenes],
 as.data.frame(t(se_spruce_VegOct_list[["V10F24-117-C1-2"]]@assays[["Spatial"]]@counts))[rownames(VegOctCorpusReps_v2$V10F24_117_C1_2$pos),combinedVegOctgenes],
 as.data.frame(t(se_spruce_VegOct_list[["V10F24-117-C1-3"]]@assays[["Spatial"]]@counts))[rownames(VegOctCorpusReps_v2$V10F24_117_C1_3$pos),combinedVegOctgenes],
 as.data.frame(t(se_spruce_VegOct_list[["V19T26-078-A1-1"]]@assays[["Spatial"]]@counts))[rownames(VegOctCorpusReps_v2$V19T26_078_A1_1$pos),combinedVegOctgenes],
 as.data.frame(t(se_spruce_VegOct_list[["V19T26-078-A1-2"]]@assays[["Spatial"]]@counts))[rownames(VegOctCorpusReps_v2$V19T26_078_A1_2$pos),combinedVegOctgenes],
 as.data.frame(t(se_spruce_VegOct_list[["V19T26-078-A1-3"]]@assays[["Spatial"]]@counts))[rownames(VegOctCorpusReps_v2$V19T26_078_A1_3$pos),combinedVegOctgenes],
 as.data.frame(t(se_spruce_VegOct_list[["V19T26-078-B1-1"]]@assays[["Spatial"]]@counts))[rownames(VegOctCorpusReps_v2$V19T26_078_B1_1$pos),combinedVegOctgenes],
 as.data.frame(t(se_spruce_VegOct_list[["V19T26-078-B1-2"]]@assays[["Spatial"]]@counts))[rownames(VegOctCorpusReps_v2$V19T26_078_B1_2$pos),combinedVegOctgenes],
  as.data.frame(t(se_spruce_VegOct_list[["V19T26-078-B1-3"]]@assays[["Spatial"]]@counts))[rownames(VegOctCorpusReps_v2$V19T26_078_B1_3$pos),combinedVegOctgenes],
 
  )

dim(combinedVegOctcorpus)
```
3319, 3340
(pixels, genes)

Check:
```{r}
head(colnames(combinedVegOctcorpus)) #gene names 
head(rownames(combinedVegOctcorpus)) #spot names
```

Save files
```{r}
saveRDS(combinedVegOctcorpus, file = paste0(directory, "combinedVegOctcorpus_20250815.rds"))
saveRDS(VegOctCorpusReps_v2, file = paste0(directory, "VegOctCorpusReplicates.rds"))
```

## Run LDA of STdeconvolve with a given K 

K=9 took 8.47 mins
K=11 13.76 mins
K=15 21.09 mins 
K=20 28.38 mins
```{r}
combinedVegOctldas_K11 <- fitLDA(counts = combinedVegOctcorpus,
                                    K = c(11),
                                    perc.rare.thresh = 0.05,
                                    seed = 0,
                                    ncores = 7,
                                    plot = TRUE)
```

Retrieve the beta and theta matrices
```{r}
optLDA_VegOct_K11 <- optimalModel(models = combinedVegOctldas_K11, opt = 11)
combinedVegOct_results_K11 <- getBetaTheta(lda = optLDA_VegOct_K11,
                        perc.filt = 0.05,
                        betaScale = 1000)
```

Save results
```{r}
write.xlsx(combinedVegOct_results_K11$theta, file = paste0(directory, "theta_matrix_VegOct_K11.xlsx"), colNames = TRUE, rowNames = TRUE)

write.xlsx(combinedVegOct_results_K11$beta, file = paste0(directory, "beta_matrix_VegOct_K11.xlsx"),colNames = TRUE, rowNames = TRUE)

#write.table(combinedVegOctgenes, file = paste0(directory_1, "combined_genes_3340_all_VegOct.txt"))
```


```{r}
saveRDS(optLDA_VegOct_K11, file = paste0(directory, "optLDA_VegOct_K11_all_samples.rds"))
saveRDS(combinedVegOctldas_K11, file = paste0(directory, "combinedVegOct_ldas_K11_all_samples_perplexity_156_numrare_2.rds"))
```

# Plot
- without H&E 

[1] "V10F24-117-C1-1" "V10F24-117-C1-2" "V10F24-117-C1-3" "V19T26-078-A1-1" "V19T26-078-A1-2" "V19T26-078-A1-3"
[7] "V19T26-078-B1-1" "V19T26-078-B1-2" "V19T26-078-B1-3"

```{r}
vizAllTopics(combinedVegOct_results_K11$theta[rownames(VegOctCorpusReps_v2$V10F24_117_C1_1$pos),], 
             VegOctCorpusReps_v2$V10F24_117_C1_1$pos,
             r = 45,
             lwd = 0.4,
             showLegend = TRUE,
             plotTitle = NA)
```

Plot for all samples without tissue image
```{r}
# "V10F24_117_C1_1", "V10F24_117_C1_2", "V10F24_117_C1_3", "V19T26_078_A1_1", "V19T26_078_A1_2", "V19T26_078_A1_3", "V19T26_078_B1_1", "V19T26_078_B1_2", "V19T26_078_B1_3"

group_cols = brewer.pal(n = 8, name = "Dark2")  # or "Set3", "Dark2", "Paired"
                                  
plt <- vizAllTopics(combinedVegOct_results_K11$theta[rownames(VegOctCorpusReps_v2$V19T26_078_B1_3$pos),], 
             VegOctCorpusReps_v2$V19T26_078_B1_3$pos,
                 group_cols = group_cols,
                 r = 45,
                 lwd = 0.4,
                 showLegend = TRUE,
                 plotTitle = NA) +
ggplot2::guides(fill=ggplot2::guide_legend(ncol=2)) +

ggplot2::theme(
  plot.background = ggplot2::element_blank()
)

plt

#Save plot
ggsave(file = paste0(directory, "plots/STdeconvolve_ALL_VegOct_11celltypes_on_V19T26_078_B1_3.pdf"), plot = plt, width = 15, height =10, dpi = 300)
```


## Plot per cell type
Plot
- without H&E 

```{r}
# "V10F24_117_C1_1", "V10F24_117_C1_2", "V10F24_117_C1_3", "V19T26_078_A1_1", "V19T26_078_A1_2", "V19T26_078_A1_3", "V19T26_078_B1_1", "V19T26_078_B1_2", "V19T26_078_B1_3"

# vizTopic() for faster plotting. Let’s visualize each deconvolved cell-type separately:
  ps <- lapply(colnames(combinedVegOct_results_K11$theta), function(celltype) {
  
  vizTopic(theta = combinedVegOct_results_K11$theta[rownames(VegOctCorpusReps_v2$V19T26_078_B1_3$pos),], 
           pos = VegOctCorpusReps_v2$V19T26_078_B1_3$pos, 
           topic = celltype, 
           plotTitle = paste0("X", celltype),
         size = 1.5, stroke = 0.5, alpha = 0.5,
         low = "white",
         high = "red")
  
})
  ps_final <- gridExtra::grid.arrange(
  grobs = ps,
  layout_matrix = rbind(c(1, 2, 3, 4),
                        c(5, 6, 7, 8),
                        c(9, 10, 11, 12))
)
  # Save vizTopic plot 
  ggsave(file =  paste0(directory, "plots/per_celltype/ALL_VegOct_K11_on_section_id_V19T26_078_B1_3_per_cell_type.pdf"), plot = ps_final, width = 15, height =12, dpi = 300)  
```


## Plot as an Assay to use tissue image
Create an assay for the STdeconvolve results to be able to use the seurat functions for plotting, to visualise the results with the tissue image. Important to distinguish deconvolved cell types for bracts vs scales 

```{r}
#se_spruce_VegOct@meta.data
se_spruce_VegOct@meta.data

theta_matrix <- as.data.frame(combinedVegOct_results_K20$theta)
#rownames(theta_matrix) <- gsub("_", "-", rownames(theta_matrix)) do not use

theta_matrix_K9 <- as.data.frame(combinedVegOct_results_K9$theta)
theta_matrix_K15 <- as.data.frame(combinedVegOct_results_K15$theta)

theta_matrix_K11 <- as.data.frame(combinedVegOct_results_K11$theta)

#Transpose it to get the spot pixels as columns, and the cell type number as rows
theta_matrix <- t(theta_matrix)
theta_matrix_K9 <- t(theta_matrix_K9)
theta_matrix_K15 <- t(theta_matrix_K15)
theta_matrix_K11 <- t(theta_matrix_K11)

se_spruce_VegOct[['STdeconvolve_K20']] <- CreateAssayObject(counts = theta_matrix)
se_spruce_VegOct[['STdeconvolve_K9']] <- CreateAssayObject(counts = theta_matrix_K9)
se_spruce_VegOct[['STdeconvolve_K15']] <- CreateAssayObject(counts = theta_matrix_K15)
se_spruce_VegOct[['STdeconvolve_K11']] <- CreateAssayObject(counts = theta_matrix_K11)
```

```{r}
saveRDS(se_spruce_VegOct, file=("/data_analysis/Spruce_STdeconvolve_Veg_Oct/spruce_compatible_obj.VegOct_all_STdecon_theta_assays.rds"))
```

```{r}
DefaultAssay(se_spruce_VegOct_assays) <- "STdeconvolve_K11"

celltypes <- rownames(se_spruce_VegOct_assays@assays$STdeconvolve_K11)

for (ct in celltypes){
  plot <- SpatialFeaturePlot(se_spruce_VegOct_assays, features = ct, alpha = 1, ncol = 3, pt.size.factor = 3) + theme(legend.position = "right")
  ggsave(file =  paste0(directory, "plots/per_celltype/tissue_overlay/tissue_VegOct_K11_cell_type_", ct, ".pdf"), plot = plot, width = 15, height =12, dpi = 300)
}

```


## Session info

<details>

<summary>

</summary>

```{r}
sessionInfo()
Sys.Date()
Sys.time()
```
