---
title: "Combining_datasets_STdeconvolve_SR"
author: "Sofia Rouot"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: html_document
format:
  html:
    code-fold: true
    toc: true
    number-sections: true
    theme: spacelab
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, warning = FALSE}
library(STdeconvolve)
library(Seurat)
library(SeuratObject)
library(openxlsx)
library(RColorBrewer)
library(ggplot2)
```


# Spruce FemOct data

Load data
```{r}
se_spruce <- readRDS(file = "spruce_compatible_obj.all_samples.rds")
```

```{r}
se_spruce_Fem <- subset(x = se_spruce, subset = stage == "Fem")
se_spruce_FemOct <- subset(x = se_spruce_Fem, subset = timepoint == "Oct")

se_spruce_FemOct_list <- SplitObject(se_spruce_FemOct, split.by = "section_id")
se_spruce_FemOct_list[[1]]@meta.data
se_spruce_FemOct_list[[8]]@meta.data
```

## Create corpus for each section id subsetted seurat object
N.B. The counts matrix are originally:
  - rows = genes 
  - columns = pixels 
  But STdeconvolve assumes the other way:
  - rows = pixels 
  - columns = genes
  
--> Use as.matrix(t(counts)) to transpose the matrix and ensure it is treated as a matrix. 

```{r}
FemOctCorpusReps_v2 <- lapply(se_spruce_FemOct_list, function(seurat_obj) {
  counts <- seurat_obj@assays[["Spatial"]]@counts
  #with transpose of matrix 
  dat <- preprocess(as.matrix(t(counts)),
                   extractPos = FALSE,
                   selected.genes = NA,
                   nTopGenes = NA,
                   genes.to.remove = NA,
                   removeAbove = 1.1,
                   removeBelow = 0.01,
                   min.reads = 1,
                   min.lib.size = 50,
                   min.detected = 1,
                   ODgenes = TRUE,
                   nTopOD = 1000,
                   od.genes.alpha = 0.05,
                   gam.k = 5) 
  dat
})
```

add the x,y positions manually in the corpus 
```{r}
unique_section_id <- unique(se_spruce_FemOct$section_id)
for (section_id in unique_section_id){
  # Access your Seurat object
  seurat_obj <- se_spruce_FemOct_list[[section_id]]
  
  # Step 1: Get the counts matrix and its pixel names
  counts <- seurat_obj@assays$Spatial@counts
  
  # Step 2: Get the true pixel coordinates (a matrix or data.frame of x/y positions)
  image_name <- names(se_spruce_FemOct_list[[section_id]]@images)
  coords <- seurat_obj@images[[image_name]]@boundaries[["centroids"]]@coords
  
  # Make sure coords rownames match column names of counts
  rownames(coords) <- colnames(counts)
  
  # Step 4: Assign coords to pos slot of corpus
  FemOctCorpusReps_v2[[section_id]]$pos <- coords

}

```


```{r}
names(FemOctCorpusReps_v2) <- c("V10F24_109_A1_1", "V10F24_109_A1_2", "V10F24_109_A1_3", "V10F24_109_B1_1", "V10F24_109_B1_2", "V10F24_109_B1_3", "V10F24_109_C1_1", "V10F24_109_C1_2")
```


```{r}
combinedFemOctgenes <- Reduce(union, list(colnames(FemOctCorpusReps_v2$V10F24_109_A1_1$corpus),
                                       colnames(FemOctCorpusReps_v2$V10F24_109_A1_2$corpus),
                                       colnames(FemOctCorpusReps_v2$V10F24_109_A1_3$corpus),
                                       colnames(FemOctCorpusReps_v2$V10F24_109_B1_1$corpus),
                                       colnames(FemOctCorpusReps_v2$V10F24_109_B1_2$corpus),
                                       colnames(FemOctCorpusReps_v2$V10F24_109_B1_3$corpus),
                                       colnames(FemOctCorpusReps_v2$V10F24_109_C1_1$corpus),
                                       colnames(FemOctCorpusReps_v2$V10F24_109_C1_2$corpus))
                           )
length(combinedFemOctgenes)
```
2765
```{r}
unique_section_id
```

```{r}
combinedFemOctcorpus <- dplyr::bind_rows(
  as.data.frame(t(se_spruce_FemOct_list[["V10F24-109-A1-1"]]@assays[["Spatial"]]@counts))[rownames(FemOctCorpusReps_v2$V10F24_109_A1_1$pos),combinedFemOctgenes],
 as.data.frame(t(se_spruce_FemOct_list[["V10F24-109-A1-2"]]@assays[["Spatial"]]@counts))[rownames(FemOctCorpusReps_v2$V10F24_109_A1_2$pos),combinedFemOctgenes],
 as.data.frame(t(se_spruce_FemOct_list[["V10F24-109-A1-3"]]@assays[["Spatial"]]@counts))[rownames(FemOctCorpusReps_v2$V10F24_109_A1_3$pos),combinedFemOctgenes],
 as.data.frame(t(se_spruce_FemOct_list[["V10F24-109-B1-1"]]@assays[["Spatial"]]@counts))[rownames(FemOctCorpusReps_v2$V10F24_109_B1_1$pos),combinedFemOctgenes],
 as.data.frame(t(se_spruce_FemOct_list[["V10F24-109-B1-2"]]@assays[["Spatial"]]@counts))[rownames(FemOctCorpusReps_v2$V10F24_109_B1_2$pos),combinedFemOctgenes],
 as.data.frame(t(se_spruce_FemOct_list[["V10F24-109-B1-3"]]@assays[["Spatial"]]@counts))[rownames(FemOctCorpusReps_v2$V10F24_109_B1_3$pos),combinedFemOctgenes],
 as.data.frame(t(se_spruce_FemOct_list[["V10F24-109-C1-1"]]@assays[["Spatial"]]@counts))[rownames(FemOctCorpusReps_v2$V10F24_109_C1_1$pos),combinedFemOctgenes],
 as.data.frame(t(se_spruce_FemOct_list[["V10F24-109-C1-2"]]@assays[["Spatial"]]@counts))[rownames(FemOctCorpusReps_v2$V10F24_109_C1_2$pos),combinedFemOctgenes],
  
  )

dim(combinedFemOctcorpus)
```
5258 2765
(pixels, genes)

```{r}
head(colnames(combinedFemOctcorpus)) #gene names 
head(rownames(combinedFemOctcorpus)) #spot names
```

Save files
```{r}
saveRDS(combinedFemOctcorpus, file = paste0(directory, "combinedFemOctcorpus.rds"))
saveRDS(FemOctCorpusReps_v2, file = paste0(directory, "FemOctCorpusReps.rds"))
```

## Run LDA
K=2 took 1.30 mins
K=9 took 12.58 mins
K=15 29.49 mins
K=20 44.29 mins
```{r}
combinedFemOctldas_K9 <- fitLDA(counts = combinedFemOctcorpus,
                                    K = c(9),
                                    perc.rare.thresh = 0.05,
                                    seed = 0,
                                    ncores = 7,
                                    plot = TRUE)
```
```{r}
optLDA_FemOct_K9 <- optimalModel(models = combinedFemOctldas_K9, opt = 9)
combinedFemOct_results_K9 <- getBetaTheta(lda = optLDA_FemOct_K9,
                        perc.filt = 0.05,
                        betaScale = 1000)
```

```{r}
write.xlsx(combinedFemOct_results_K20$theta, file = paste0(directory, "theta_matrix_FemOct_K20.xlsx"), colNames = TRUE, rowNames = TRUE)

write.xlsx(combinedFemOct_results_K20$beta, file = paste0(directory, "beta_matrix_FemOct_K20.xlsx"),colNames = TRUE, rowNames = TRUE)

#write.table(combinedFemOctgenes, file = paste0(directory_1, "combined_genes_all_FemOct.txt"))

saveRDS(optLDA_FemOct_K20, file = paste0(directory, "optlda_K20_FemOct_all_femoct_samples.rds"))
saveRDS(combinedFemOctldas_K20, file = paste0(directory, "lda_K20_FemOct_all_femoct_samples_perplexity__numrare_.rds"))
```

# Plot
- without H&E 
```{r}
vizAllTopics(combinedFemOct_results_K20$theta[rownames(FemOctCorpusReps_v2$V10F24_109_A1_2$pos),], 
             FemOctCorpusReps_v2$V10F24_109_A1_2$pos,
             r = 45,
             lwd = 0.4,
             showLegend = TRUE,
             plotTitle = NA)
```

Plot for all samples without tissue image
```{r}
# "V10F24_109_A1_1", "V10F24_109_A1_2", "V10F24_109_A1_3", "V10F24_109_B1_1", "V10F24_109_B1_2", "V10F24_109_B1_3", "V10F24_109_C1_1", "V10F24_109_C1_2"
group_cols = brewer.pal(n = 8, name = "Dark2")  # or "Set3", "Dark2", "Paired"
                                  
plt <- vizAllTopics(combinedFemOct_results_K20$theta[rownames(FemOctCorpusReps_v2$V10F24_109_C1_2$pos),], 
             FemOctCorpusReps_v2$V10F24_109_C1_2$pos,
                 group_cols = group_cols,
                 r = 45,
                 lwd = 0.4,
                 showLegend = TRUE,
                 plotTitle = NA) +
ggplot2::guides(fill=ggplot2::guide_legend(ncol=2)) +


ggplot2::theme(
  plot.background = ggplot2::element_blank()
)
plt
#Save plot
ggsave(file = paste0(directory, "plots/STdeconvolve_ALL_FemOct_20celltypes_on_V10F24_109_C1_2.pdf"), plot = plt, width = 15, height =10, dpi = 300)
  

                                  
plt <- vizAllTopics(theta_K9[rownames(FemOctCorpusReps_v2$V10F24_109_C1_2$pos),], 
             FemOctCorpusReps_v2$V10F24_109_C1_2$pos,
                 group_cols = group_cols,
                 r = 45,
                 lwd = 0.4,
                 showLegend = TRUE,
                 plotTitle = NA) +
ggplot2::guides(fill=ggplot2::guide_legend(ncol=2)) +


ggplot2::theme(
  plot.background = ggplot2::element_blank()
)
plt
#Save plot
ggsave(file = paste0(directory, "STdeconvolve_ALL_FemOct_9celltypes_on_V10F24_109_C1_2.pdf"), plot = plt, width = 15, height =10, dpi = 300)
```


## Plot per cell type
Plot
- without H&E 

```{r}
# vizTopic() for faster plotting. Let’s visualize each deconvolved cell-type separately:
  ps <- lapply(colnames(combinedFemOct_results_K9$theta), function(celltype) {
  
  vizTopic(theta = combinedFemOct_results_K9$theta[rownames(FemOctCorpusReps_v2$V10F24_109_C1_2$pos),], 
           pos = FemOctCorpusReps_v2$V10F24_109_C1_2$pos, 
           topic = celltype, 
           plotTitle = paste0("X", celltype),
         size = 1.5, stroke = 0.5, alpha = 0.5,
         low = "white",
         high = "red")
  
})
  ps_final <- gridExtra::grid.arrange(
  grobs = ps,
  layout_matrix = rbind(c(1, 2, 3, 4),
                        c(5, 6, 7, 8),
                        c(9, 10, 11, 12))
)
  # Save vizTopic plot 
  ggsave(file =  paste0(directory, "plots/ALL_FemOct_K9_on_section_id_V10F24_109_C1_2_per_cell_type.pdf"), plot = ps_final, width = 15, height =12, dpi = 300)  
```



## Plot as an Assay to use tissue image
Create an assay for the STdeconvolve results to be able to use the seurat functions for plotting, to visualise the results with the tissue image. Important to distinguish deconvolved cell types for bracts vs scales 

Load object and the theta matrices 
```{r}

FemOct_theta_K9 <- read.xlsx("/data_analysis/Spruce_STdeconvolve_Fem_Oct/results_K_9/all_FemOct_samples/theta_matrix_FemOct.xlsx", colNames = TRUE, rowNames = TRUE)

FemOct_theta_K15 <- read.xlsx("/data_analysis/Spruce_STdeconvolve_Fem_Oct/results_K_celltype_15/all_FemOct_samples/theta_matrix_FemOct.xlsx", colNames = TRUE, rowNames = TRUE)

FemOct_theta_K20 <- read.xlsx("/data_analysis/Spruce_STdeconvolve_Fem_Oct/results_K_20/all_FemOct_samples/theta_matrix_FemOct_K20.xlsx", colNames = TRUE, rowNames = TRUE)
```

### Create Assay for each STdeconvolve theta matrix (K=9, 15, 20)
```{r}
se_spruce_FemOct@meta.data
unique(se_spruce_FemOct$section_id) #8

theta_matrix_K9 <- as.data.frame(FemOct_theta_K9)
theta_matrix_K15 <- as.data.frame(FemOct_theta_K15)
theta_matrix_K20 <- as.data.frame(FemOct_theta_K20)

#Transpose it to get the spot pixels as columns, and the cell type number as rows
theta_matrix_K9 <- t(theta_matrix_K9)
theta_matrix_K15 <- t(theta_matrix_K15)
theta_matrix_K20 <- t(theta_matrix_K20)

se_spruce_FemOct[['STdeconvolve_K9']] <- CreateAssayObject(counts = theta_matrix_K9)
se_spruce_FemOct[['STdeconvolve_K15']] <- CreateAssayObject(counts = theta_matrix_K15)
se_spruce_FemOct[['STdeconvolve_K20']] <- CreateAssayObject(counts = theta_matrix_K20)

```

```{r}
saveRDS(se_spruce_FemOct, file=("/data_analysis/Spruce_STdeconvolve_Fem_Oct/spruce_compatible_obj.FemOct_all_STdecon_theta_assays.rds"))
```

### Visualize the STdeconvolve cell types with the tissue 
```{r}
DefaultAssay(se_spruce_FemOct) <- "STdeconvolve_K9"

celltypes <- rownames(se_spruce_FemOct@assays$STdeconvolve_K9)

for (ct in celltypes){
  plot <- SpatialFeaturePlot(se_spruce_FemOct, features = ct, alpha = 1, ncol = 3, pt.size.factor = 3) + theme(legend.position = "right")
  ggsave(file =  paste0(directory, "plots/per_cell_type_tissue_overlay/tissue_FemOct_K9_cell_type_", ct, ".pdf"), plot = plot, width = 15, height =12, dpi = 300)
}

```


## Session info

<details>

<summary>

</summary>

```{r}
sessionInfo()
Sys.Date()
Sys.time()
```
