---
title: "Combining_datasets_STdeconvolve"
author: "Sofia Rouot"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: html_document
format:
  html:
    code-fold: true
    toc: true
    number-sections: true
    theme: spacelab
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, warning = FALSE}
library(STdeconvolve)
library(Seurat)
library(SeuratObject)
library(openxlsx)
library(RColorBrewer)
library(ggplot2)
```


# Spruce AcroOct data

Load data
```{r}
se_spruce <- readRDS(file = "spruce_compatible_obj.all_samples.rds")
```

```{r}
se_spruce_Acro <- subset(x = se_spruce, subset = stage == "Acro")
se_spruce_AcroOct <- subset(x = se_spruce_Acro, subset = timepoint == "Oct")

se_spruce_AcroOct_list <- SplitObject(se_spruce_AcroOct, split.by = "section_id")
se_spruce_AcroOct_list[[1]]@meta.data
se_spruce_AcroOct_list[[7]]@meta.data
```

```{r}
unique_section_id <- unique(se_spruce_AcroOct$section_id)
unique_section_id
```

## Create corpus for each section id subsetted seurat object
N.B. The counts matrix are originally:
  - rows = genes 
  - columns = pixels 
  But STdeconvolve assumes the other way:
  - rows = pixels 
  - columns = genes
  
--> Use as.matrix(t(counts)) to transpose the matrix and ensure it is treated as a matrix. 
```{r}
AcroOctCorpusReps <- lapply(se_spruce_AcroOct_list, function(seurat_obj) {
  counts <- seurat_obj@assays[["Spatial"]]@counts
  #with transpose of matrix 
  dat <- preprocess(as.matrix(t(counts)),
                   extractPos = FALSE, #do it manually in next step
                   selected.genes = NA,
                   nTopGenes = NA,
                   genes.to.remove = NA,
                   removeAbove = 1.1, # keep genes in up to 110% of pixels (i.e. no upper filter)
                   removeBelow = 0.01, # only remove genes in <0.1% of spots
                   min.reads = 1, #removes genes (rows) expressed in <1 pixels
                   min.lib.size = 50, #removes pixels (columns) with <50 total counts
                   min.detected = 1,
                   ODgenes = TRUE,
                   nTopOD = 1000, # allow max 1000 overdispersed genes
                   od.genes.alpha = 0.05,
                   gam.k = 5) 
  dat
})
```

Add the x,y positions manually in the corpus 
```{r}
unique_section_id <- unique(se_spruce_AcroOct$section_id)
for (section_id in unique_section_id){
  # Access your Seurat object
  seurat_obj <- se_spruce_AcroOct_list[[section_id]]
  
  # Step 1: Get the counts matrix and its pixel names
  counts <- seurat_obj@assays$Spatial@counts
  
  # Step 2: Get the true pixel coordinates (a matrix or data.frame of x/y positions)
  image_name <- names(se_spruce_AcroOct_list[[section_id]]@images)
  coords <- seurat_obj@images[[image_name]]@boundaries[["centroids"]]@coords
  
  # Make sure coords rownames match column names of counts
  rownames(coords) <- colnames(counts)
  
  # Step 4: Assign coords to pos slot of corpus
  AcroOctCorpusReps[[section_id]]$pos <- coords

}

```


```{r}
names(AcroOctCorpusReps) <- c("V10F24_117_D1_1", "V10F24_117_D1_2", "V10F24_117_D1_3", "V19T26_078_C1_1", "V19T26_078_C1_2", "V19T26_078_D1_1", "V19T26_078_D1_2")
```

## Get the combined significant genes from all corpuses
```{r}
combinedAcroOctgenes <- Reduce(union, list(colnames(AcroOctCorpusReps$V10F24_117_D1_1$corpus),
                                           colnames(AcroOctCorpusReps$V10F24_117_D1_2$corpus),
                                           colnames(AcroOctCorpusReps$V10F24_117_D1_3$corpus),
                                           colnames(AcroOctCorpusReps$V19T26_078_C1_1$corpus),
                                           colnames(AcroOctCorpusReps$V19T26_078_C1_2$corpus),
                                           colnames(AcroOctCorpusReps$V19T26_078_D1_1$corpus),
                                           colnames(AcroOctCorpusReps$V19T26_078_D1_2$corpus))
                           )
length(combinedAcroOctgenes)
```
3066
```{r}
unique_section_id
```

## Bind each count matrix into a common corpus with only the combined genes
```{r}
combinedAcroOctcorpus <- dplyr::bind_rows(
  as.data.frame(t(se_spruce_AcroOct_list[["V10F24-117-D1-1"]]@assays[["Spatial"]]@counts))[rownames(AcroOctCorpusReps$V10F24_117_D1_1$pos),combinedAcroOctgenes],
 as.data.frame(t(se_spruce_AcroOct_list[["V10F24-117-D1-2"]]@assays[["Spatial"]]@counts))[rownames(AcroOctCorpusReps$V10F24_117_D1_2$pos),combinedAcroOctgenes],
 as.data.frame(t(se_spruce_AcroOct_list[["V10F24-117-D1-3"]]@assays[["Spatial"]]@counts))[rownames(AcroOctCorpusReps$V10F24_117_D1_3$pos),combinedAcroOctgenes],
 as.data.frame(t(se_spruce_AcroOct_list[["V19T26-078-C1-1"]]@assays[["Spatial"]]@counts))[rownames(AcroOctCorpusReps$V19T26_078_C1_1$pos),combinedAcroOctgenes],
 as.data.frame(t(se_spruce_AcroOct_list[["V19T26-078-C1-2"]]@assays[["Spatial"]]@counts))[rownames(AcroOctCorpusReps$V19T26_078_C1_2$pos),combinedAcroOctgenes],
 as.data.frame(t(se_spruce_AcroOct_list[["V19T26-078-D1-1"]]@assays[["Spatial"]]@counts))[rownames(AcroOctCorpusReps$V19T26_078_D1_1$pos),combinedAcroOctgenes],
 as.data.frame(t(se_spruce_AcroOct_list[["V19T26-078-D1-2"]]@assays[["Spatial"]]@counts))[rownames(AcroOctCorpusReps$V19T26_078_D1_2$pos),combinedAcroOctgenes]
  )

dim(combinedAcroOctcorpus)
```
3399 3066
(pixels, genes)

```{r}
head(colnames(combinedAcroOctcorpus)) #gene names 
head(rownames(combinedAcroOctcorpus)) #spot names
```

Save files
```{r}
saveRDS(combinedAcroOctcorpus, file = paste0(directory, "combinedAcroOctcorpus.rds"))
saveRDS(AcroOctCorpusReps, file = paste0(directory_1, "AcroOctCorpusReps.rds"))

write.table(combinedAcroOctgenes, file = paste0(directory, "combined_genes_all_AcroOct.txt"))
```

## Run the LDA deconvolution
K=9 took 9.63 mins 
K=15 took 20.51 mins 
K=20 took 28.32 mins
```{r}
combinedAcroOctldas_K20 <- fitLDA(counts = combinedAcroOctcorpus,
                                    K = c(20),
                                    perc.rare.thresh = 0.05,
                                    seed = 0,
                                    ncores = 7,
                                    plot = TRUE)
```

```{r}
optLDA_Acro_15 <- optimalModel(models = combinedAcroOctldas, opt = 15)
combinedAcroOct_results_K15 <- getBetaTheta(lda = optLDA_Acro_15,
                        perc.filt = 0.05,
                        betaScale = 1000)

optLDA_Acro_9 <- optimalModel(models = combinedAcroOctldas_K9, opt = 9)
combinedAcroOct_results_K9 <- getBetaTheta(lda = optLDA_Acro_9,
                        perc.filt = 0.05,
                        betaScale = 1000)

optLDA_Acro_20 <- optimalModel(models = combinedAcroOctldas_K20, opt = 20)
combinedAcroOct_results_K20 <- getBetaTheta(lda = optLDA_Acro_20,
                        perc.filt = 0.05,
                        betaScale = 1000)
```

Save files
```{r}
write.xlsx(combinedAcroOct_results_K20$theta, file = paste0(directory, "theta_matrix_AcroOct_K20.xlsx"), colNames = TRUE, rowNames = TRUE)

write.xlsx(combinedAcroOct_results_K20$beta, file = paste0(directory, "beta_matrix_AroOct_K20.xlsx"),colNames = TRUE, rowNames = TRUE)

saveRDS(optLDA_Acro_20, file = paste0(directory, "lda_K20_AcroOct_all_samples.rds"))
saveRDS(combinedAcroOctldas_K20, file = paste0(directory, "lda_K20_AcroOct_all_samples_perplexity_74_numRare_20.rds"))
```


# Plot
- without H&E 

```{r}
# "V10F24_117_D1_1", "V10F24_117_D1_2", "V10F24_117_D1_3", "V19T26_078_C1_1", "V19T26_078_C1_2", "V19T26_078_D1_1", "V19T26_078_D1_2"
vizAllTopics(combinedAcroOct_results_K20$theta[rownames(AcroOctCorpusReps$V10F24_117_D1_1$pos),], 
             AcroOctCorpusReps$V10F24_117_D1_1$pos,
             r = 45,
             lwd = 0.4,
             showLegend = TRUE,
             plotTitle = NA)
```

Do it on all AcroOct samples/sections
```{r}
# "V10F24_117_D1_1", "V10F24_117_D1_2", "V10F24_117_D1_3", "V19T26_078_C1_1", "V19T26_078_C1_2", "V19T26_078_D1_1", "V19T26_078_D1_2"

group_cols = brewer.pal(n = 8, name = "Dark2")  # or "Set3", "Dark2", "Paired"
                                  
plt <- vizAllTopics(combinedAcroOct_results_K20$theta[rownames(AcroOctCorpusReps$V19T26_078_D1_2$pos),], 
             AcroOctCorpusReps$V19T26_078_D1_2$pos,
                 group_cols = group_cols,
                 r = 45,
                 lwd = 0.4,
                 showLegend = TRUE,
                 plotTitle = NA) +
ggplot2::guides(fill=ggplot2::guide_legend(ncol=2)) +


ggplot2::theme(
  plot.background = ggplot2::element_blank()
)
#plt
#Save plot
ggsave(file = paste0(directory, "plots/STdeconvolve_ALL_AcroOct_20celltypes_on_V19T26_078_D1_2.pdf"), plot = plt, width = 15, height =10, dpi = 300)
```


## Plot per cell type
Plot
- without H&E 

```{r}
# "V10F24_117_D1_1", "V10F24_117_D1_2", "V10F24_117_D1_3", "V19T26_078_C1_1", "V19T26_078_C1_2", "V19T26_078_D1_1", "V19T26_078_D1_2"
# vizTopic() for faster plotting. Let’s visualize each deconvolved cell-type separately:
  
  ps <- lapply(colnames(combinedAcroOct_results_K20$theta), function(celltype) {
  
  vizTopic(theta = combinedAcroOct_results_K20$theta[rownames(AcroOctCorpusReps$V19T26_078_D1_2$pos),], 
           pos = AcroOctCorpusReps$V19T26_078_D1_2$pos, 
           topic = celltype, 
           plotTitle = paste0("X", celltype),
         size = 1.5, stroke = 0.5, alpha = 0.5,
         low = "white",
         high = "red")
  
})
  ps_final <- gridExtra::grid.arrange(
  grobs = ps,
  layout_matrix = rbind(c(1, 2, 3, 4),
                        c(5, 6, 7, 8),
                        c(9, 10, 11, 12),
                        c(13,14,15,16),
                        c(17,18,19,20))
)
  # Save vizTopic plot 
  ggsave(file =  paste0(directory, "plots/ALL_FemOct_K20_on_section_id_V19T26_078_D1_2_per_cell_type.pdf"), plot = ps_final, width = 15, height =12, dpi = 300)  
```



## Plot as an Assay to use tissue image
Create an assay for the STdeconvolve results to be able to use the seurat functions for plotting, to visualise the results with the tissue image. Important to distinguish deconvolved cell types for bracts vs scales 

Load object and the theta matrices 
```{r}

AcroOct_theta_K9 <- read.xlsx("/data_analysis/Spruce_STdeconvolve_Acr_Oct/results/K_9/all_AcroOct_samples/theta_matrix_AcroOct_K9.xlsx", colNames = TRUE, rowNames = TRUE)

AcroOct_theta_K15 <- read.xlsx("/data_analysis/Spruce_STdeconvolve_Acr_Oct/results/K_15/all_AcroOct_samples/theta_matrix_AcroOct_K15.xlsx", colNames = TRUE, rowNames = TRUE)

AcroOct_theta_K20 <- read.xlsx("/data_analysis/Spruce_STdeconvolve_Acr_Oct/results/results_K_20/all_AcroOct_samples/theta_matrix_AcroOct_K20.xlsx", colNames = TRUE, rowNames = TRUE)
```

### Create Assay for each STdeconvolve theta matrix (K=9, 15, 20)
```{r}
se_spruce_AcroOct@meta.data
unique(se_spruce_AcroOct$section_id) #7

theta_matrix_K9 <- as.data.frame(AcroOct_theta_K9)
theta_matrix_K15 <- as.data.frame(AcroOct_theta_K15)
theta_matrix_K20 <- as.data.frame(AcroOct_theta_K20)

#Transpose it to get the spot pixels as columns, and the cell type number as rows
theta_matrix_K9 <- t(theta_matrix_K9)
theta_matrix_K15 <- t(theta_matrix_K15)
theta_matrix_K20 <- t(theta_matrix_K20)

se_spruce_AcroOct[['STdeconvolve_K9']] <- CreateAssayObject(counts = theta_matrix_K9)
se_spruce_AcroOct[['STdeconvolve_K15']] <- CreateAssayObject(counts = theta_matrix_K15)
se_spruce_AcroOct[['STdeconvolve_K20']] <- CreateAssayObject(counts = theta_matrix_K20)

```

```{r}
saveRDS(se_spruce_AcroOct, file=("/data_analysis/Spruce_STdeconvolve_Acr_Oct/spruce_compatible_obj.AcroOct_all_STdecon_theta_assays.rds"))
```

### Visualize the STdeconvolve cell types with the tissue 
```{r}
DefaultAssay(se_spruce_AcroOct) <- "STdeconvolve_K20"

celltypes <- rownames(se_spruce_AcroOct@assays$STdeconvolve_K20)

for (ct in celltypes){
  plot <- SpatialFeaturePlot(se_spruce_AcroOct, features = ct, alpha = 1, ncol = 3, pt.size.factor = 3) + theme(legend.position = "right")
  ggsave(file =  paste0(directory, "plots/per_cell_type_tissue_overlay/tissue_FemOct_K20_cell_type_", ct, ".pdf"), plot = plot, width = 15, height =12, dpi = 300)
}

```


## Session info

<details>

<summary>

</summary>

```{r}
sessionInfo()
Sys.Date()
Sys.time()
```
