---
title: "STdeconvolve_Results_Interpretation_Spruce_SR"
author: "Sofia Rouot"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: html_document
format:
  html:
    code-fold: true
    toc: true
    number-sections: true
    theme: spacelab
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, warning = FALSE}
library(STdeconvolve)
library(Seurat)
library(SeuratObject)
library(openxlsx)
library(RColorBrewer)
library(ggplot2)
library(ggridges)
library(tidyverse)
library(dplyr)
library(readxl)
library(pheatmap)
library(grid)
```


### Load input data (deconProp/beta matrix from K=15 for FemOct)

FemOct
```{r}
FemOct_beta_K15 <- read.xlsx("/data_analysis/Spruce_STdeconvolve_Fem_Oct/results_K_celltype_15/all_FemOct_samples/beta_matrix_FemOct.xlsx", colNames = TRUE, rowNames = TRUE)
```

AcroOct
```{r}
AcroOct_beta_K20 <- read.xlsx("/data_analysis/Spruce_STdeconvolve_Acr_Oct/results/results_K_20/all_AcroOct_samples/beta_matrix_AroOct_K20.xlsx", colNames = TRUE, rowNames = TRUE)

AcroOct_theta_K20 <- read.xlsx("data_analysis/Spruce_STdeconvolve_Acr_Oct/results/results_K_20/all_AcroOct_samples/theta_matrix_AcroOct_K20.xlsx", colNames = TRUE, rowNames = TRUE)
```

VegOct 
```{r}
VegOct_beta_K <- read.xlsx("/data_analysis/Spruce_STdeconvolve_Veg_Oct/K11/all_VegOct_samples/beta_matrix_VegOct_K11.xlsx", colNames = TRUE, rowNames = TRUE)
```

### 1. Extract the top gene markers from each cell type into a dataframe 

Latest version with Log2FC for Fem, Veg, Acro
LOG2FC > 0.25
```{r}
#Fem
deconGexp <- data.frame(FemOct_beta_K15)
# Ensure matrix is numeric
deconGexp <- as.matrix(sapply(deconGexp, as.numeric))
rownames(deconGexp) <- 1:nrow(deconGexp)

#Acro
deconGexp <- data.frame(AcroOct_beta_K20)
# Ensure matrix is numeric
deconGexp <- as.matrix(sapply(deconGexp, as.numeric))
rownames(deconGexp) <- 1:nrow(deconGexp)

#Veg
deconGexp <- data.frame(VegOct_beta_K)
deconGexp <- as.matrix(sapply(deconGexp, as.numeric))
rownames(deconGexp) <- 1:nrow(deconGexp)


top_markers_A <- lapply(1:nrow(deconGexp), function(cell_idx) {
  
  # Select genes highly expressed in this cell type
  highgexp_idx <- which(deconGexp[cell_idx, ] > 0.5)
  
  if (length(highgexp_idx) == 0) return(NULL)  # skip if none pass threshold
  
  # Get gene names for those indices
  highgexp_genes <- colnames(deconGexp)[highgexp_idx]
  
  ## high log2(fold-change) compared to other deconvolved cell-types
  log2fc <- log2(
    deconGexp[cell_idx, highgexp_idx, drop=TRUE] /
      colMeans(deconGexp[-cell_idx, highgexp_idx, drop=FALSE])
  )
  names(log2fc) <- highgexp_genes
  
    # Keep only  log2FC > 0.25
  log2fc <- log2fc[log2fc > 0.25]
  if (length(log2fc) == 0) return(NA)  # skip if none are positive
  
  # Sort and take top 50
  log2fc_sorted <- sort(log2fc, decreasing=TRUE)
  top50 <- head(log2fc_sorted, 50)
  
  # Return as data frame with ranking
  data.frame(
    cell_type = rownames(deconGexp)[cell_idx],
    gene = names(log2fc_sorted), #top50 or log2fc_sorted
    log2FC = as.numeric(log2fc_sorted), #top50 or log2fc_sorted
    rank = seq_along(log2fc_sorted), #top50 or log2fc_sorted
    stringsAsFactors = FALSE
  )
})

# Combine all into one data frame
top_markers_A_df <- bind_rows(top_markers_A)
```


```{r}
#Save
directory <- "/data_analysis/Results_Interpretation/marker_genes_all/"
write.xlsx(top_markers_Fem_df, file = paste0(directory, "FemOct_K15_all_gene_markers_log2FC.xlsx"),colNames = TRUE, rowNames = TRUE)

write.xlsx(top_markers_Veg_df, file = paste0(directory, "VegOct_K11_all_gene_markers_log2FC.xlsx"),colNames = TRUE, rowNames = TRUE)

write.xlsx(top_markers_A_df, file = paste0(directory, "AcroOct_K20_all_gene_markers_log2FC.xlsx"),colNames = TRUE, rowNames = TRUE)

```

Plot log2fc distributions per cell type
```{r}
# Check the actual labels
#unique(top_markers_Fem_df$cell_type)

cell_order <- c(1:20)
top_markers_df$cell_type <- factor(top_markers_df$cell_type, levels = cell_order)

p_vln_box <- ggplot(top_markers_df, aes(x = cell_type, y = log2FC, fill = cell_type)) +
  geom_violin(trim = FALSE, alpha = 0.5) +
  geom_boxplot(width = 0.1, outlier.shape = NA) +
  theme_minimal(base_size = 14) +
  scale_y_continuous(breaks = seq(0, max(top_markers_df$log2FC), by = 1)) +
  labs(
    title = "Distribution of log2FC values per AcroOct cell type (violin + boxplot)",
    x = "Cell type",
    y = "log2FC"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

p_vln_box

p_ridge <- ggplot(top_markers_df, aes(x = log2FC, y = cell_type, fill = cell_type)) +
  geom_density_ridges(alpha = 0.7, scale = 1.2) +
  theme_minimal(base_size = 14) +
  scale_x_continuous(breaks = seq(0, max(top_markers_df$log2FC), by = 1)) +
  labs(
    title = "Density of log2FC values per AcroOct cell type (ridge)",
    x = "log2FC",
    y = "Cell type"
  ) +
  theme(legend.position = "none")

p_ridge

p_final <- p_vln_box / p_ridge
p_final

directory <- "/data_analysis/Results_Interpretation/top50_gene_markers_per_celltype/log2fc_distribution_plots/"
ggsave(file = paste0(directory, "plot_log2FC_distribution_top50_gene_markers_per_AcroOct_celltypes.pdf"), plot = p_final, width = 15, height =10, dpi = 300)

ggplot(top_markers_df, aes(x = cell_type, y = log2FC, fill = cell_type)) +
  geom_boxplot(outlier.alpha = 0.3) +
  facet_wrap(~cell_type, scales = "free_x") +
  theme_minimal(base_size = 14) +
  labs(
    title = "Distribution of log2FC values per cell type (top 50 markers)",
    x = "Cell type",
    y = "log2FC"
  ) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none",
    strip.text = element_text(size = 12)
  )

```


### 2. Compute overlap between Fem15 and Acro20
Between each cell type of the top markers of FemOct and AcroOct, compute overlap of similar genes 

Read top markers files if needed
```{r}
sheets <- excel_sheets("/data_analysis/Results_Interpretation/top50_gene_markers_per_celltype/FemOct_K15_top50_gene_markers.xlsx")

# Read each sheet into a list
top_markers_FemOct_list <- lapply(sheets, function(sheet) {
  read.xlsx("/data_analysis/Results_Interpretation/top50_gene_markers_per_celltype/FemOct_K15_top50_gene_markers.xlsx", colNames = FALSE, sheet = sheet)
})

# Name the list by sheet names
names(top_markers_FemOct_list) <- sheets



sheets <- excel_sheets("/data_analysis/Results_Interpretation/top50_gene_markers_per_celltype/AcroOct_K20_top50_gene_markers.xlsx")

## For Acro
# Read each sheet into a list
top_markers_AcroOct_list <- lapply(sheets, function(sheet) {
  read.xlsx("/data_analysis/Results_Interpretation/top50_gene_markers_per_celltype/AcroOct_K20_top50_gene_markers.xlsx", colNames = FALSE, sheet = sheet)
})

# Name the list by sheet names
names(top_markers_AcroOct_list) <- sheets


## For Veg
sheets <- excel_sheets("/data_analysis/Results_Interpretation/top50_gene_markers_per_celltype/VegOct_K11_top50_gene_markers.xlsx")
# Read each sheet into a list
top_markers_VegOct_list <- lapply(sheets, function(sheet) {
  read.xlsx("/data_analysis/Results_Interpretation/top50_gene_markers_per_celltype/VegOct_K11_top50_gene_markers.xlsx", colNames = FALSE, sheet = sheet)
})

# Name the list by sheet names
names(top_markers_VegOct_list) <- sheets

```

For high overlap, visualize those cell types to see if spatial expression is similar as well.

### 3. Compute overlap between Fem15 and Veg11 (heatmap)

Add annotations to the heatmap 
```{r}

top_markers_FemOct <- as.data.frame(top_markers_FemOct_list)
#names(top_markers_FemOct) <- paste0("Fem_CellType_", 1:15)
top_markers_VegOct <- as.data.frame(top_markers_VegOct_list)
#names(top_markers_VegOct) <- paste0("Veg_CellType_", 1:11)
#top_markers_AcroOct <- as.data.frame(top_markers_AcroOct_list)
names(top_markers_AcroOct) <- paste0("Acro_CellType_", 1:20)

#Add cell type annotation 
FemOct_annot <- c("Fem_CellType_1"="Fem_1_Pith bottom", 
                  "Fem_CellType_2"="Fem_2_Vasculature", 
                  "Fem_CellType_3"="Fem_3_LOs-bracts",
                  "Fem_CellType_4"="Fem_4_Meristem top",
                  "Fem_CellType_5"="Fem_5_LOs-scales",
                  "Fem_CellType_6"="Fem_6_LOs towards vasculature",
                  "Fem_CellType_7"="Fem_7_Vasculature towards LOs-scales",
                  "Fem_CellType_8"="Fem_8_LO-scales",
                  "Fem_CellType_9"="Fem_9_LOs-vasculature-towards scales ",
                  "Fem_CellType_10"="Fem_10_LOs-scales",
                  "Fem_CellType_11"="Fem_11_Pith middle",
                  "Fem_CellType_12"="Fem_12_Pith middle top",
                  "Fem_CellType_13"="Fem_13_Meristem-pith-vasculature and LOs",
                  "Fem_CellType_14"="Fem_14_Meristem-scales with meristematic cells",
                  "Fem_CellType_15"="Fem_15_LOs bracts")

names(top_markers_FemOct) <- FemOct_annot[names(top_markers_FemOct)]

VegOct_annot <- c("Veg_CellType_1"="Veg_1_Meristem-top LOs", 
                  "Veg_CellType_2"="Veg_2_Vasculature-LOs", 
                  "Veg_CellType_3"="Veg_3_Pith bottom-lowest LOs",
                  "Veg_CellType_4"="Veg_4_LOs",
                  "Veg_CellType_5"="Veg_5_Pith bottom",
                  "Veg_CellType_6"="Veg_6_LOs bottom",
                  "Veg_CellType_7"="Veg_7_LOs top",
                  "Veg_CellType_8"="Veg_8_Vasculature bottom",
                  "Veg_CellType_9"="Veg_9_Pith middle",
                  "Veg_CellType_10"="Veg_10_LOs-vasculature",
                  "Veg_CellType_11"="Veg_11_Pith top")
names(top_markers_VegOct) <- VegOct_annot[names(top_markers_VegOct)]


AcroOct_annot <- c("Acro_CellType_1"="Acro_1_Vasculature-LOs", 
                  "Acro_CellType_2"="Acro_2_Pith", 
                  "Acro_CellType_3"="Acro_3_Vasculature-LOs top",
                  "Acro_CellType_4"="Acro_4_LOs bottom",
                  "Acro_CellType_5"="Acro_5_LOs top-scales",
                  "Acro_CellType_6"="Acro_6_Bottom pith-LOs",
                  "Acro_CellType_7"="Acro_7_Pith bottom",
                  "Acro_CellType_8"="Acro_8_Meristem-vasculature and scales",
                  "Acro_CellType_9"="Acro_9_Pith bottom middle",
                  "Acro_CellType_10"="Acro_10_Meristem-scales",
                  "Acro_CellType_11"="Acro_11_LOs top",
                  "Acro_CellType_12"="Acro_12_LOs",
                  "Acro_CellType_13"="Acro_13_Pith-LOs",
                  "Acro_CellType_14"="Acro_14_Vasculature-LOs bottom",
                  "Acro_CellType_15"="Acro_15_Pith top",
                  "Acro_CellType_16"="Acro_16_LOs bottom",
                  "Acro_CellType_17"="Acro_17_Pith top middle",
                  "Acro_CellType_18"="Acro_18_LOs top middle (transition zone)-vaculature-scales",
                  "Acro_CellType_19"="Acro_19_LOs top middle (transition zone)-bracts",
                  "Acro_CellType_20"="Acro_20_Pith top middle")
names(top_markers_AcroOct) <- AcroOct_annot[names(top_markers_AcroOct)]
```


```{r}
#Fem
top_markers_Fem_list <- lapply(top_markers_Fem, function(df) df$gene)
names(top_markers_Fem_list) <- paste0("Fem_CellType_", 1:15)
str(top_markers_Fem_list)
names(top_markers_Fem_list) <- FemOct_annot[names(top_markers_Fem_list)]
top_markers_FemOct <- top_markers_Fem_list

#Acro
top_markers_Acro_list <- lapply(top_markers_Acro, function(df) df$gene)
names(top_markers_Acro_list) <- paste0("Acro_CellType_", 1:20)
str(top_markers_Acro_list)
names(top_markers_Acro_list) <- AcroOct_annot[names(top_markers_Acro_list)]
top_markers_AcroOct <- top_markers_Acro_list

#Veg
top_markers_Veg_list <- lapply(top_markers_Veg, function(df) df$gene)
names(top_markers_Veg_list) <- paste0("Veg_CellType_", 1:11)
str(top_markers_Veg_list)
names(top_markers_Veg_list) <- VegOct_annot[names(top_markers_Veg_list)]
top_markers_VegOct <- top_markers_Veg_list
```


```{r}
# Compute overlap matrix (15 × 20)
overlap_matrix <- matrix(0, 
                         nrow = length(top_markers_FemOct), 
                         ncol = length(top_markers_VegOct),
                         dimnames = list(names(top_markers_FemOct), 
                                         names(top_markers_VegOct)))

for (f_ct in names(top_markers_FemOct)) {
  for (a_ct in names(top_markers_VegOct)) {
    
    common_genes <- intersect(top_markers_FemOct[[f_ct]], 
                              top_markers_VegOct[[a_ct]])
    
    overlap_matrix[f_ct, a_ct] <- length(common_genes) / 50 * 100
  }
}

write.xlsx(overlap_matrix, file = paste0(directory, "heatmaps/heatmap_overlap_Fem_Veg/overlap_matrix_top50_gene_markers_FemOct_vs_VegOct.xlsx"),colNames = TRUE, rowNames = TRUE)
```


Group them by main cell types in heatmap
```{r}
annotation_map_Fem <- c(
  "Fem_1_Pith bottom" = "Pith", 
  "Fem_2_Vasculature" = "Vasculature", 
  "Fem_3_LOs-bracts"= "Lateral Organs",
  "Fem_4_Meristem top"= "Meristem",
  "Fem_5_LOs-scales"= "Lateral Organs",
  "Fem_6_LOs towards vasculature"= "Lateral Organs",
  "Fem_7_Vasculature towards LOs-scales"= "Vasculature",
  "Fem_8_LO-scales"= "Lateral Organs",
  "Fem_9_LOs-vasculature-towards scales "= "Lateral Organs",
  "Fem_10_LOs-scales"= "Lateral Organs",
  "Fem_11_Pith middle"= "Pith",
  "Fem_12_Pith middle top"= "Pith",
  "Fem_13_Meristem-pith-vasculature and LOs"= "Meristem",
  "Fem_14_Meristem-scales with meristematic cells"= "Meristem",
  "Fem_15_LOs bracts"= "Lateral Organs"
)

cell_classes_Fem <- annotation_map_Fem[rownames(overlap_matrix)]

annotation_map_Veg <- c(
  "Veg_1_Meristem-top LOs" = "Meristem", 
  "Veg_2_Vasculature-LOs"= "Vasculature", 
  "Veg_3_Pith bottom-lowest LOs"= "Pith",
  "Veg_4_LOs"= "Lateral Organs",
  "Veg_5_Pith bottom"= "Pith",
  "Veg_6_LOs bottom"= "Lateral Organs",
  "Veg_7_LOs top"= "Lateral Organs",
  "Veg_8_Vasculature bottom"= "Vasculature",
  "Veg_9_Pith middle"= "Pith",
  "Veg_10_LOs-vasculature"= "Lateral Organs",
  "Veg_11_Pith top"= "Pith"
)

cell_classes_Veg <- annotation_map_Veg[colnames(overlap_matrix)]

row_order <- order(cell_classes_Fem)
col_order <- order(cell_classes_Veg)
```

Re-order matrix based on these annotations
```{r}
overlap_matrix_ordered <- overlap_matrix[row_order, col_order]
```

```{r}
# Heatmap visualization
p <- pheatmap(overlap_matrix_ordered,
         annotation_row = data.frame(Class = cell_classes_Fem),
         annotation_col = data.frame(Class = cell_classes_Veg),
         cluster_rows = FALSE,
         cluster_cols = FALSE,   # cluster Veg cell types
         display_numbers = TRUE, # show % values
         number_format = "%.1f", # one decimal
         fontsize_number = 7,
         fontsize_row = 7,
         fontsize_col =7,
         main = "Gene Marker Overlap Fem15 vs Veg11 (%)")

ggsave(file = paste0(directory, "heatmaps/heatmap_overlap_Fem_Veg/heatmap_overlap_matrix_gene_markers_FemOct_vs_VegOct_ordered.pdf"), plot = p, width = 15, height =10, dpi = 300)
```


### 4. Compute overlap between Acro20 and Veg11 
```{r}
#top_markers_AcroOct <- as.data.frame(top_markers_AcroOct_list)
#names(top_markers_AcroOct) <- paste0("Acro_CellType_", 1:20)
#top_markers_VegOct <- as.data.frame(top_markers_Veg)
#names(top_markers_VegOct) <- paste0("Veg_CellType_", 1:11)

# Compute overlap matrix
overlap_matrix <- matrix(0, 
                         nrow = length(top_markers_AcroOct), 
                         ncol = length(top_markers_VegOct),
                         dimnames = list(names(top_markers_AcroOct), 
                                         names(top_markers_VegOct)))

for (f_ct in names(top_markers_AcroOct)) {
  for (a_ct in names(top_markers_VegOct)) {
    
    common_genes <- intersect(top_markers_AcroOct[[f_ct]], 
                              top_markers_VegOct[[a_ct]])
    
    overlap_matrix[f_ct, a_ct] <- length(common_genes) / 50 * 100
  }
}

write.xlsx(overlap_matrix, file = paste0(directory, "heatmaps/heatmap_overlap_Acro_Veg/overlap_matrix_top50_gene_markers_AcroOct_vs_VegOct.xlsx"),colNames = TRUE, rowNames = TRUE)
```

```{r}
cell_classes_Acro <- annotation_map_Acro[rownames(overlap_matrix)]
cell_classes_Veg <- annotation_map_Veg[colnames(overlap_matrix)]

row_order <- order(cell_classes_Acro)
col_order <- order(cell_classes_Veg)

overlap_matrix_ordered <- overlap_matrix[row_order, col_order]
```


```{r}
# Heatmap visualization
p <- pheatmap(overlap_matrix_ordered,
         annotation_row = data.frame(Class = cell_classes_Acro),
         annotation_col = data.frame(Class = cell_classes_Veg),
         cluster_rows = FALSE,
         cluster_cols = FALSE,   # cluster Veg cell types
         display_numbers = TRUE, # show % values
         number_format = "%.1f", # one decimal
         fontsize_number = 7,
         fontsize_row = 7,
         fontsize_col =7,
         main = "Gene Marker Overlap Acro20 vs Veg11 (%)")

ggsave(file = paste0(directory, "heatmaps/heatmap_overlap_Acro_Veg/heatmap_overlap_matrix_gene_markers_AcroOct_vs_VegOct_ordered.pdf"), plot = p, width = 15, height =10, dpi = 300)
```


### 5. Compute overlap between Fem15 and Acro20
```{r}
#top_markers_AcroOct <- as.data.frame(top_markers_AcroOct_list)
#names(top_markers_AcroOct) <- paste0("Acro_CellType_", 1:20)
#top_markers_VegOct <- as.data.frame(top_markers_Veg)
#names(top_markers_VegOct) <- paste0("Veg_CellType_", 1:11)

# Compute overlap matrix
overlap_matrix <- matrix(0, 
                         nrow = length(top_markers_FemOct), 
                         ncol = length(top_markers_AcroOct),
                         dimnames = list(names(top_markers_FemOct), 
                                         names(top_markers_AcroOct)))

for (f_ct in names(top_markers_FemOct)) {
  for (a_ct in names(top_markers_AcroOct)) {
    
    common_genes <- intersect(top_markers_FemOct[[f_ct]], 
                              top_markers_AcroOct[[a_ct]])
    
    overlap_matrix[f_ct, a_ct] <- length(common_genes) / 50 * 100
  }
}

write.xlsx(overlap_matrix, file = paste0(directory, "heatmaps/heatmap_overlap_Fem_Acro/overlap_matrix_top50_gene_markers_FemOct_vs_AcroOct.xlsx"),colNames = TRUE, rowNames = TRUE)
```

Group them by main cell types in heatmap
```{r}
annotation_map_Fem <- c(
  "Fem_1_Pith bottom" = "Pith", 
  "Fem_2_Vasculature" = "Vasculature", 
  "Fem_3_LOs-bracts"= "Lateral Organs",
  "Fem_4_Meristem top"= "Meristem",
  "Fem_5_LOs-scales"= "Lateral Organs",
  "Fem_6_LOs towards vasculature"= "Lateral Organs",
  "Fem_7_Vasculature towards LOs-scales"= "Vasculature",
  "Fem_8_LO-scales"= "Lateral Organs",
  "Fem_9_LOs-vasculature-towards scales "= "Lateral Organs",
  "Fem_10_LOs-scales"= "Lateral Organs",
  "Fem_11_Pith middle"= "Pith",
  "Fem_12_Pith middle top"= "Pith",
  "Fem_13_Meristem-pith-vasculature and LOs"= "Meristem",
  "Fem_14_Meristem-scales with meristematic cells"= "Meristem",
  "Fem_15_LOs bracts"= "Lateral Organs"
)

cell_classes_Fem <- annotation_map_Fem[rownames(overlap_matrix)]

annotation_map_Acro <- c(
  "Acro_1_Vasculature-LOs" = "Vasculature", 
  "Acro_2_Pith" = "Pith", 
  "Acro_3_Vasculature-LOs top" = "Vasculature",
  "Acro_4_LOs bottom" = "Lateral Organs",
  "Acro_5_LOs top-scales" = "Lateral Organs",
  "Acro_6_Bottom pith-LOs" = "Pith",
  "Acro_7_Pith bottom" = "Pith",
  "Acro_8_Meristem-vasculature and scales" = "Meristem",
  "Acro_9_Pith bottom middle" = "Pith",
  "Acro_10_Meristem-scales" = "Meristem",
  "Acro_11_LOs top" = "Lateral Organs",
  "Acro_12_LOs" = "Lateral Organs",
  "Acro_13_Pith-LOs" = "Pith",
  "Acro_14_Vasculature-LOs bottom" = "Vasculature",
  "Acro_15_Pith top" = "Pith",
  "Acro_16_LOs bottom" = "Lateral Organs",
  "Acro_17_Pith top middle" = "Pith",
  "Acro_18_LOs top middle (transition zone)-vaculature-scales" = "Lateral Organs",
  "Acro_19_LOs top middle (transition zone)-bracts" = "Lateral Organs",
  "Acro_20_Pith top middle" = "Pith"
)

cell_classes_Acro <- annotation_map_Acro[colnames(overlap_matrix)]

row_order <- order(cell_classes_Fem)
col_order <- order(cell_classes_Acro)
```

Re-order matrix based on these annotations
```{r}
overlap_matrix_ordered <- overlap_matrix[row_order, col_order]
```

```{r}
# Heatmap visualization
p <- pheatmap(overlap_matrix_ordered,
         annotation_row = data.frame(Class = cell_classes_Fem),
         annotation_col = data.frame(Class = cell_classes_Acro),
         cluster_rows = FALSE,
         cluster_cols = FALSE,   # cluster Acro cell types
         display_numbers = TRUE, # show % values
         number_format = "%.1f", # one decimal
         fontsize_number = 7,
         fontsize_row = 7,
         fontsize_col =7,
         main = "Gene Marker Overlap Fem15 vs Acro20 (%)")

ggsave(file = paste0(directory, "heatmaps/heatmap_overlap_Fem_Acro/heatmap_overlap_matrix_top50_gene_markers_FemOct_vs_AcroOct_ordered.pdf"), plot = p, width = 15, height =10, dpi = 300)
```



### 6. Compute overlap within Fem 15 cell types
```{r}
#top_markers_FemOct <- as.data.frame(top_markers_FemOct_list)
#names(top_markers_FemOct) <- paste0("Fem_CellType_", 1:15)


# Compute overlap matrix (15 × 20)
overlap_matrix_FemFem <- matrix(0, 
                         nrow = length(top_markers_FemOct), 
                         ncol = length(top_markers_FemOct),
                         dimnames = list(names(top_markers_FemOct), 
                                         names(top_markers_FemOct)))

for (f_ct in names(top_markers_FemOct)) {
  for (a_ct in names(top_markers_FemOct)) {
    
    common_genes <- intersect(top_markers_FemOct[[f_ct]], 
                              top_markers_FemOct[[a_ct]])
    
    overlap_matrix_FemFem[f_ct, a_ct] <- length(common_genes) / 50 * 100
  }
}

write.xlsx(overlap_matrix_FemFem, file = paste0(directory, "heatmaps/heatmap_overlap_Fem_Fem/overlap_matrix_top50_gene_markers_FemOct_vs_FemOct.xlsx"),colNames = TRUE, rowNames = TRUE)
```

```{r}
cell_classes_Fem <- annotation_map_Fem[rownames(overlap_matrix_FemFem)]
cell_classes_Fem2 <- annotation_map_Fem[colnames(overlap_matrix_FemFem)]

row_order <- order(cell_classes_Fem)
col_order <- order(cell_classes_Fem)

overlap_matrix_ordered <- overlap_matrix_FemFem[row_order, col_order]
```

```{r}
# Heatmap visualization
p <- pheatmap(overlap_matrix_ordered,
         annotation_row = data.frame(Class = cell_classes_Fem),
         annotation_col = data.frame(Class = cell_classes_Fem),
         cluster_rows = FALSE,
         cluster_cols = FALSE,   # cluster Female cell types
         display_numbers = TRUE, # show % values
         number_format = "%.1f", # one decimal
         fontsize_number = 7,
         fontsize_row = 7,
         fontsize_col =7,
         main = "Gene Marker Overlap within FemOct15 (%)")
#grid.text("FemOct Cell Types K15", x = 0.5, y = 0.01, gp = gpar(fontsize = 10))  # x-axis
#grid.text("FemOct Cell Types K15", x = 0.98, y = 0.5, rot = 90, gp = gpar(fontsize = 10))

ggsave(file = paste0(directory, "heatmaps/heatmap_overlap_Fem_Fem/heatmap_overlap_matrix_top50_gene_markers_FemOct_vs_FemOct_ordered.pdf"), plot = p, width = 15, height =10, dpi = 300)
```

### 7. Compute overlap within Acro 20 cell types
```{r}
overlap_matrix <- matrix(0, 
                         nrow = length(top_markers_AcroOct), 
                         ncol = length(top_markers_AcroOct),
                         dimnames = list(names(top_markers_AcroOct), 
                                         names(top_markers_AcroOct)))

for (f_ct in names(top_markers_AcroOct)) {
  for (a_ct in names(top_markers_AcroOct)) {
    
    common_genes <- intersect(top_markers_AcroOct[[f_ct]], 
                              top_markers_AcroOct[[a_ct]])
    
    overlap_matrix[f_ct, a_ct] <- length(common_genes) / 50 * 100
  }
}

write.xlsx(overlap_matrix, file = paste0(directory, "heatmaps/heatmap_overlap_Acro_Acro/overlap_matrix_top50_gene_markers_AcroOct_vs_AcroOct.xlsx"),colNames = TRUE, rowNames = TRUE)
```

```{r}
cell_classes_Acro <- annotation_map_Acro[rownames(overlap_matrix)]

row_order <- order(cell_classes_Acro)
col_order <- order(cell_classes_Acro)

overlap_matrix_ordered <- overlap_matrix[row_order, col_order]
```

```{r}
# Heatmap visualization
p <- pheatmap(overlap_matrix_ordered,
         annotation_row = data.frame(Class = cell_classes_Acro),
         annotation_col = data.frame(Class = cell_classes_Acro),
         cluster_rows = FALSE,
         cluster_cols = FALSE,   # cluster Acro cell types
         display_numbers = TRUE, # show % values
         number_format = "%.1f", # one decimal
         fontsize_number = 7,
         fontsize_row = 7,
         fontsize_col =7,
         main = "Gene Marker Overlap within Acro20 (%)")

ggsave(file = paste0(directory, "heatmaps/heatmap_overlap_Acro_Acro/heatmap_overlap_matrix_top50_gene_markers_AcroOct_vs_AcroOct_ordered.pdf"), plot = p, width = 15, height =10, dpi = 300)
```

### 8. Compute overlap within Veg 11 cell types
```{r}
overlap_matrix <- matrix(0, 
                         nrow = length(top_markers_VegOct), 
                         ncol = length(top_markers_VegOct),
                         dimnames = list(names(top_markers_VegOct), 
                                         names(top_markers_VegOct)))

for (f_ct in names(top_markers_VegOct)) {
  for (a_ct in names(top_markers_VegOct)) {
    
    common_genes <- intersect(top_markers_VegOct[[f_ct]], 
                              top_markers_VegOct[[a_ct]])
    
    overlap_matrix[f_ct, a_ct] <- length(common_genes) / 50 * 100
  }
}

write.xlsx(overlap_matrix, file = paste0(directory, "heatmaps/heatmap_overlap_Veg_Veg/overlap_matrix_top50_gene_markers_VegOct_vs_VegOct.xlsx"),colNames = TRUE, rowNames = TRUE)
```

```{r}
cell_classes_Veg <- annotation_map_Veg[rownames(overlap_matrix)]

row_order <- order(cell_classes_Veg)
col_order <- order(cell_classes_Veg)

overlap_matrix_ordered <- overlap_matrix[row_order, col_order]
```

```{r}
# Heatmap visualization
p <- pheatmap(overlap_matrix_ordered,
         annotation_row = data.frame(Class = cell_classes_Veg),
         annotation_col = data.frame(Class = cell_classes_Veg),
         cluster_rows = FALSE,
         cluster_cols = FALSE,   # cluster Veg cell types
         display_numbers = TRUE, # show % values
         number_format = "%.1f", # one decimal
         fontsize_number = 7,
         fontsize_row = 7,
         fontsize_col =7,
         main = "Gene Marker Overlap within Veg11 (%)")

ggsave(file = paste0(directory, "heatmaps/heatmap_overlap_Veg_Veg/heatmap_overlap_matrix_top50_gene_markers_VegOct_vs_VegOct_ordered.pdf"), plot = p, width = 15, height =10, dpi = 300)
```


# 9. GO term analysis for marker genes

### a. Get file with Arabidopsis annotations
Read in the annotations files. spruce gene --> arabidopsis gene  
```{r}
FemOct_Arabidopsis_annotations <- read.xlsx("TableSX_STdeconvolve_GO_summary_VegOct_FemOct_AcroOct.xlsx", colNames = TRUE, rowNames = TRUE) # take sheet 2

AcroOct_Arabidopsis_annotations <- read.xlsx("TableSX_STdeconvolve_GO_summary_VegOct_FemOct_AcroOct.xlsx", colNames = TRUE, rowNames = TRUE) #take sheet 3

VegOct_Arabidopsis_annotations <- read.xlsx("TableSX_STdeconvolve_GO_summary_VegOct_FemOct_AcroOct.xlsx", colNames = TRUE, rowNames = TRUE) #take sheet 1
```

Filter for log2FC >1. 
```{r}
FemOct_Arabidopsis_annotations_db <- as.data.frame(FemOct_Arabidopsis_annotations)
AcroOct_Arabidopsis_annotations_db <- as.data.frame(AcroOct_Arabidopsis_annotations)
VegOct_Arabidopsis_annotations_db <- as.data.frame(VegOct_Arabidopsis_annotations)
```

```{r}
FemOct_Arabidopsis_annotations_db <- FemOct_Arabidopsis_annotations_db %>%
  dplyr::filter(log2FC > 1)

AcroOct_Arabidopsis_annotations_db <- AcroOct_Arabidopsis_annotations_db %>%
  dplyr::filter(log2FC > 1)

VegOct_Arabidopsis_annotations_db <- VegOct_Arabidopsis_annotations_db %>%
  dplyr::filter(log2FC > 1)
```

GO enrichment with Arabidopsis IDs column "ATG.Best.DIAMOND"

### b. GO analysis for each bud type: FemOct, AcroOct, VegOct

For this, we are using clusterProfiler and org.At.tair.db.

```{r}
#renv::install("clusterProfiler")
library(clusterProfiler)
library(org.At.tair.db)
```

### For FemOct

```{r}
library(dplyr)
library(stringr)

# FemOct_Arabidopsis_annotations_db$ATG.Best.DIAMOND
# FemOct_Arabidopsis_annotations_db$cell_type
# Make sure the CellType column is a factor or character
FemOct_Arabidopsis_annotations_db$cell_type <- as.factor(FemOct_Arabidopsis_annotations_db$cell_type)

# List of unique cell types
cell_types <- unique(FemOct_Arabidopsis_annotations_db$cell_type)

# Create a list to store all cell type results
all_cell_results <- list()

# Loop over each cell type
for (cell in cell_types) {
  
  # Subset genes for this cell type
  genes <- FemOct_Arabidopsis_annotations_db %>%
    filter(cell_type == cell) %>%
    pull(ATG.Best.DIAMOND) %>%
    na.omit() %>%
    unique()
  
  # Skip if no genes
  if (length(genes) == 0) next
  
  # GO enrichment
  ego <- enrichGO(
    gene          = genes,
    OrgDb         = org.At.tair.db,
    keyType       = "TAIR",
    ont           = "BP",         # Change to "BP" MF" or "CC" if needed
    pAdjustMethod = "BH",
    pvalueCutoff  = 0.05,
    qvalueCutoff  = 0.2,
    readable      = FALSE
  )
  
  # Skip if no enrichment
  if (is.null(ego) || nrow(ego) == 0) next
  
  # Dotplot
  p <- dotplot(ego, showCategory = 20, orderBy = "p.adjust") +
    ggtitle(paste("Top 20 GO Terms for FemOct cell type", cell))
  
  # Save plot
  ggsave(filename = paste0("/GO_terms/FemOct/GO_dotplots_top20_BP_terms/Top20_GO_dotplot_celltype", cell, ".png"), plot = p, width = 8, height = 12)
  
  
   # Extract enrichment results correctly
  ego_df <- ego@result
  
  # Filter Count column for at least 2 genes
  ego_df <- ego_df %>%
    filter(Count >= 2)
  
  # Add cell type column
  ego_df$CellType <- cell
  
  # Select key columns
  ego_summary <- ego_df %>%
    dplyr::select(CellType,
           ID, Description,
           Count,
           pvalue, p.adjust, qvalue,
           geneID) 
  
  # Save per cell type
  write_csv(ego_summary, paste0("/GO_terms/FemOct/GO_summary_BP_tables/GO_summary_celltype_", cell, ".csv"))
  
  all_cell_results[[cell]] <- ego_summary
  
  
  print(paste("Finished GO analysis for cell type:", cell))
}

# Combine all into one table
combined_GO_summary <- bind_rows(all_cell_results)

# Save combined table
write_csv(combined_GO_summary, "/GO_terms/FemOct/GO_summary_BP_tables/All_cells_GO_summary_FemOct.csv")

```

### For AcroOct
```{r}
# Make sure the CellType column is a factor or character
AcroOct_Arabidopsis_annotations_db$cell_type <- as.factor(AcroOct_Arabidopsis_annotations_db$cell_type)

# List of unique cell types
cell_types <- unique(AcroOct_Arabidopsis_annotations_db$cell_type)

# Create a list to store all cell type results
all_cell_results <- list()

# Loop over each cell type
for (cell in cell_types) {
  
  # Subset genes for this cell type
  genes <- AcroOct_Arabidopsis_annotations_db %>%
    filter(cell_type == cell) %>%
    pull(ATG.Best.DIAMOND) %>%
    na.omit() %>%
    unique()
  
  # Skip if no genes
  if (length(genes) == 0) next
  
  # GO enrichment
  ego <- enrichGO(
    gene          = genes,
    OrgDb         = org.At.tair.db,
    keyType       = "TAIR",
    ont           = "BP",         # Change to "MF" or "CC" if needed
    pAdjustMethod = "BH",
    pvalueCutoff  = 0.05,
    qvalueCutoff  = 0.2,
    readable      = FALSE,
    minGSSize = 2 
  )
  
  # Skip if no enrichment
  if (is.null(ego) || nrow(ego) == 0) next
  
  # Dotplot
  p <- dotplot(ego, showCategory = 20, orderBy = "p.adjust") +
    ggtitle(paste("Top 20 GO Terms for cell type", cell))
  
  # Save plot
  ggsave(filename = paste0("/GO_terms/AcroOct/GO_dotplots_top20_BP_terms/Top20_GO_dotplot_celltype_", cell, ".png"), plot = p, width = 8, height = 12)
  
  
   # Extract enrichment results correctly
  ego_df <- ego@result
  
  # Filter Count column for at least 2 genes
  ego_df <- ego_df %>%
    filter(Count >= 2)
  
  # Add cell type column
  ego_df$CellType <- cell
  
  # Select key columns
  ego_summary <- ego_df %>%
    dplyr::select(CellType,
           ID, Description,
           Count,
           pvalue, p.adjust, qvalue,
           geneID) 
  
  # Save per cell type
  write_csv(ego_summary, paste0("/GO_terms/AcroOct/GO_summary_BP_tables/GO_summary_celltype_", cell, ".csv"))
  
  all_cell_results[[cell]] <- ego_summary
  
  print(paste("Finished GO analysis for cell type:", cell))
}

# Combine all into one table
combined_GO_summary <- bind_rows(all_cell_results)

# Save combined table
write_csv(combined_GO_summary, "/GO_terms/AcroOct/GO_summary_BP_tables_new_names/All_cells_GO_summary_AcroOct.csv")
```

### For VegOct
```{r}
# Make sure the CellType column is a factor or character
VegOct_Arabidopsis_annotations_db$cell_type <- as.factor(VegOct_Arabidopsis_annotations_db$cell_type)

# List of unique cell types
cell_types <- unique(VegOct_Arabidopsis_annotations_db$cell_type)

# Create a list to store all cell type results
all_cell_results <- list()

# Loop over each cell type
for (cell in cell_types) {
  
  # Subset genes for this cell type
  genes <- VegOct_Arabidopsis_annotations_db %>%
    filter(cell_type == cell) %>%
    pull(ATG.Best.DIAMOND) %>%
    na.omit() %>%
    unique()
  
  # Skip if no genes
  if (length(genes) == 0) next
  
  # GO enrichment
  ego <- enrichGO(
    gene          = genes,
    OrgDb         = org.At.tair.db,
    keyType       = "TAIR",
    ont           = "BP",         # Change to "MF" or "CC" if needed
    pAdjustMethod = "BH",
    pvalueCutoff  = 0.05,
    qvalueCutoff  = 0.2,
    readable      = FALSE,
    minGSSize = 2 
  )
  
  # Skip if no enrichment
  if (is.null(ego) || nrow(ego) == 0) next
  
  # Dotplot
  p <- dotplot(ego, showCategory = 20, orderBy = "p.adjust") +
    ggtitle(paste("Top 20 GO Terms for AcroOct cell type", cell))
  
  # Save plot
  ggsave(filename = paste0("/GO_terms/VegOct/GO_dotplots_top20_BP_terms/Top20_GO_dotplot_celltype_", cell, ".png"), plot = p, width = 8, height = 12)
  
  
   # Extract enrichment results correctly
  ego_df <- ego@result
  
  # Filter Count column for at least 2 genes
  #ego_df <- ego_df %>%
  #  filter(Count >= 2)
  
  # Add cell type column
  ego_df$CellType <- cell
  
  # Select key columns
  ego_summary <- ego_df %>%
    dplyr::select(CellType,
           ID, Description,
           Count,
           pvalue, p.adjust, qvalue,
           geneID) 
  
  # Save per cell type
  write_csv(ego_summary, paste0("/GO_terms/VegOct/GO_summary_BP_tables/all_genes_GO_summary_celltype_", cell, ".csv"))
  
  all_cell_results[[cell]] <- ego_summary
  
  print(paste("Finished GO analysis for cell type:", cell))
}

# Combine all into one table
combined_GO_summary <- bind_rows(all_cell_results)

# Save combined table
write_csv(combined_GO_summary, "/GO_terms/VegOct/GO_summary_BP_tables_new_names/All_cells_GO_summary_VegOct.csv")
```


# 10. Pearson correlations between cell types
Between cell types: 
  Fem vs Acro, 
  Fem vs Veg, 
  Acro vs Veg, 
  Fem vs Fem, 
  Acro vs Acro, 
  Veg vs Veg

Load all 3 beta matrices
FemOct
```{r}
FemOct_beta_K15 <- read.xlsx("/beta_matrix_FemOct_K15.xlsx", colNames = TRUE, rowNames = TRUE)
```

AcroOct
```{r}
AcroOct_beta_K20 <- read.xlsx("beta_matrix_AroOct_K20.xlsx", colNames = TRUE, rowNames = TRUE)

```

VegOct 
```{r}
VegOct_beta_K11 <- read.xlsx("beta_matrix_VegOct_K11.xlsx", colNames = TRUE, rowNames = TRUE)
```

```{r}
deconGexp_Fem <- data.frame(FemOct_beta_K15)
# Ensure matrix is numeric
deconGexp_Fem <- as.matrix(sapply(deconGexp_Fem, as.numeric))
rownames(deconGexp_Fem) <- 1:nrow(deconGexp_Fem)

deconGexp_AcroOct <- data.frame(AcroOct_beta_K20)
# Ensure matrix is numeric
deconGexp_AcroOct <- as.matrix(sapply(deconGexp_AcroOct, as.numeric))
rownames(deconGexp_AcroOct) <- 1:nrow(deconGexp_AcroOct)

deconGexp_Veg <- data.frame(VegOct_beta_K11)
# Ensure matrix is numeric
deconGexp_Veg <- as.matrix(sapply(deconGexp_Veg, as.numeric))
rownames(deconGexp_Veg) <- 1:nrow(deconGexp_Veg)
```

```{r}
m1 <- deconGexp_Fem
m2 <- deconGexp_AcroOct

rownames(m1) <- paste0("X", rownames(m1))
rownames(m2) <- paste0("X", rownames(m2))

corMtx <- STdeconvolve::getCorrMtx(m1 = as.matrix(m1),
                                   m2 = as.matrix(m2),
                                   type = "b")
```

```{r}
#pairs <- STdeconvolve::lsatPairs(corMtx) 
```
N.B. the commented line goes from nxm to nxn if m>n. For ex: from 11x20 to 11x11 for Veg vs Acro, so some Acro cell types are removed.
It does an optimal assignment:
1 => 3, 2 => 18, 3 => 6, 4 => 12, 5 => 20, 6 => 8, 7 => 10, 8 => 19, 9 => 9, 10 => 5, 11 => 15. So those not here are excluded. 
Therefore, we will skip this optimal assignment step


```{r}
plt <- STdeconvolve::correlationPlot(mat = corMtx,
                                     colLabs = "FemOct_K15", # aka x-axis, and rows of matrix
                                     rowLabs = "AcroOct_K20", # aka y-axis, and columns of matrix
                                     title = "Transcriptional profile correlation within VegOct deconvolved cell types",
                                     annotation = TRUE) +
                    ggplot2::theme(axis.text.x = ggplot2::element_text(size=12, color = "black", hjust = 0, vjust = 0.5),
                                   axis.text.y = ggplot2::element_text(size=12, color = "black"),
                                   axis.title.y = ggplot2::element_text(size=15),
                                   axis.title.x = ggplot2::element_text(size=15),
                                   plot.title = ggplot2::element_text(size=0),
                                   legend.text = ggplot2::element_text(size = 15, colour = "black"),
                                   legend.title = ggplot2::element_text(size = 15, colour = "black", angle = 90),
                                   panel.background = ggplot2::element_blank(),
                                   ## border around plot
                                   panel.border = ggplot2::element_rect(fill = NA, color = "black", size = 2),
                                   plot.background = ggplot2::element_blank()
                                   # legend.position="none"
                                   )

plt
```

```{r}
ggsave(filename = paste0("Pearson_corr_cell_types_FemOctK15_AcroOctK20_full", ".pdf"), plot = plt, width = 12, height = 16)
```



## Session info

<details>

<summary>

</summary>

```{r}
sessionInfo()
Sys.Date()
Sys.time()
```
