---
title: "Load Data to Seurat, and filtering"
author: "Yuvarani Masarapu"
date: "4/9/2021"
output: html_document
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.path='figs/', fig.width=14, fig.height=10, warning=FALSE, message=FALSE, echo = FALSE)
```

```{r libraries-load, echo=FALSE}
suppressPackageStartupMessages({
  library(STutility)
  library(Seurat)
  library(harmony)
  library(gridExtra)
  library(pals)
  library(akima)
  library(readr)
  library(plotly)
  library(tibble)
  library(raster)
  library(dplyr)
  library(stringr)
  library(magrittr)
  library(ggplot2)
  library(imager)
  library(Matrix)
  library(cowplot)
  require(data.table)
  library(magick)
  library(grid)
  library(SeuratObject)
})
```

## Extract raw data as data frame
```{r}
#raw_data_ST <- data.frame()
samples <- c("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide1_V19T26-100_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide1_V19T26-100_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide1_V19T26-100_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide1_V19T26-100_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide1_V19T26-100_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide1_V19T26-100_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide1_V19T26-100_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide1_V19T26-100_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide1_V19T26-100_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide1_V19T26-100_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide1_V19T26-100_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide2_V19T26-029_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide2_V19T26-029_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide2_V19T26-029_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide2_V19T26-029_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide2_V19T26-029_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide2_V19T26-029_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide2_V19T26-029_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide2_V19T26-029_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide2_V19T26-029_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide2_V19T26-029_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide2_V19T26-029_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide2_V19T26-029_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide3_V19T26-078_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide3_V19T26-078_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide3_V19T26-078_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide3_V19T26-078_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide3_V19T26-078_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide3_V19T26-078_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide3_V19T26-078_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide3_V19T26-078_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide3_V19T26-078_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide3_V19T26-078_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide4_V19T26-079_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide5_V10F24-116_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide5_V10F24-116_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide5_V10F24-116_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide5_V10F24-116_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide5_V10F24-116_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide6_V10F24-117_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide6_V10F24-117_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide6_V10F24-117_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide6_V10F24-117_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide6_V10F24-117_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide6_V10F24-117_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide6_V10F24-117_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide6_V10F24-117_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide6_V10F24-117_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide6_V10F24-117_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide6_V10F24-117_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide6_V10F24-117_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide6_V10F24-117_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide6_V10F24-117_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide7_V10F24-109_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide7_V10F24-109_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide7_V10F24-109_A1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide7_V10F24-109_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide7_V10F24-109_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide7_V10F24-109_B1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide7_V10F24-109_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide7_V10F24-109_C1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide7_V10F24-109_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide7_V10F24-109_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide7_V10F24-109_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide7_V10F24-109_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide7_V10F24-109_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide7_V10F24-109_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide7_V10F24-109_D1_trimmed_stdata.tsv",
            "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/gene_counts/Slide7_V10F24-109_D1_trimmed_stdata.tsv")


spotfiles <- list.files(path = "/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/spot_files_from_json", pattern = ".tsv$", full.names = T)

section_id <- c("V19T26-100-A1-1", "V19T26-100-A1-2", "V19T26-100-B1-1", "V19T26-100-B1-2", "V19T26-100-B1-3", "V19T26-100-C1-1", "V19T26-100-C1-2", "V19T26-100-C1-3", "V19T26-100-D1-1", "V19T26-100-D1-2", "V19T26-100-D1-3", "V19T26-029-A1-1", "V19T26-029-A1-2", "V19T26-029-B1-1", "V19T26-029-B1-2", "V19T26-029-C1-1", "V19T26-029-C1-2", "V19T26-029-C1-3", "V19T26-029-C1-4", "V19T26-029-D1-1", "V19T26-029-D1-2", "V19T26-029-D1-3", "V19T26-029-D1-4", "V19T26-078-A1-1", "V19T26-078-A1-2", "V19T26-078-A1-3", "V19T26-078-B1-1", "V19T26-078-B1-2", "V19T26-078-B1-3", "V19T26-078-C1-1", "V19T26-078-C1-2", "V19T26-078-D1-1", "V19T26-078-D1-2", "V19T26-079-A1-1", "V19T26-079-A1-2", "V19T26-079-A1-3", "V19T26-079-A1-4", "V19T26-079-A1-5", "V19T26-079-B1-1", "V19T26-079-B1-2", "V19T26-079-B1-3", "V19T26-079-B1-4", "V19T26-079-C1-1", "V19T26-079-C1-2", "V19T26-079-C1-3", "V19T26-079-C1-4", "V19T26-079-C1-5", "V19T26-079-D1-1", "V19T26-079-D1-2", "V19T26-079-D1-3", "V19T26-079-D1-4", "V19T26-079-D1-5", "V19T26-079-D1-6", "V10F24-116-B1-1", "V10F24-116-B1-2", "V10F24-116-B1-3", "V10F24-116-B1-4", "V10F24-116-B1-5", "V10F24-117-A1-1", "V10F24-117-A1-2", "V10F24-117-A1-3", "V10F24-117-A1-4", "V10F24-117-B1-1", "V10F24-117-B1-2", "V10F24-117-B1-3", "V10F24-117-B1-4", "V10F24-117-C1-1", "V10F24-117-C1-2", "V10F24-117-C1-3", "V10F24-117-D1-1", "V10F24-117-D1-2", "V10F24-117-D1-3", "V10F24-109-A1-1", "V10F24-109-A1-2", "V10F24-109-A1-3", "V10F24-109-B1-1", "V10F24-109-B1-2", "V10F24-109-B1-3", "V10F24-109-C1-1", "V10F24-109-C1-2", "V10F24-109-D1-1", "V10F24-109-D1-2", "V10F24-109-D1-3", "V10F24-109-D1-4", "V10F24-109-D1-5", "V10F24-109-D1-6", "V10F24-109-D1-7", "V10F24-109-D1-8")

array_id <- c(rep("V19T26-100-A1", 2), rep("V19T26-100-B1",3) , rep("V19T26-100-C1",3), rep("V19T26-100-D1", 3), rep("V19T26-029-A1",2), rep("V19T26-029-B1", 2), rep("V19T26-029-C1",4), rep("V19T26-029-D1", 4), rep("V19T26-078-A1",3), rep("V19T26-078-B1", 3), rep("V19T26-078-C1",2), rep("V19T26-078-D1", 2), rep("V19T26-079-A1",5), rep("V19T26-079-B1", 4), rep("V19T26-079-C1",5), rep("V19T26-079-D1", 6), rep("V10F24-116-B1",5), rep("V10F24-117-A1", 4), rep("V10F24-117-B1",4), rep("V10F24-117-C1", 3), rep("V10F24-117-D1",3), rep("V10F24-109-A1", 3), rep("V10F24-109-B1",3), rep("V10F24-109-C1", 2), rep("V10F24-109-D1",8))

slide_id <- c(rep("V19T26-100", 11), rep("V19T26-029", 12), rep("V19T26-078", 10), rep("V19T26-079", 20), rep("V10F24-116",5), rep("V10F24-117", 14), rep("V10F24-109",16))


timepoint<- c(rep("Sept",2),rep("Aug",9), rep("Sept",4), rep("Aug",4), rep("Sept",4), rep("Oct",10), rep("Sept",14), rep("Aug",15), rep("Sept",4), rep("Oct",14), rep("Aug",4), rep("Sept",4))

stage<- c(rep("Fem",15), rep("Acro",8), rep("Veg",6), rep("Acro",13), rep("Veg",16), rep("Acro",4), rep("Veg",7), rep("Acro",3), rep("Fem",8), rep("Veg",8))

bud<- c(rep("SeptFem1",2), rep("AugFem1",3), rep("AugFem2",6), rep("SeptFem2",2), rep("SeptFem3",2), rep("AugAcro1",2), rep("AugAcro2",2), rep("SeptAcro1",4), rep("OctVeg1",3), rep("OctVeg2",3), rep("OctAcro1",2), rep("OctAcro2",2), rep("SeptAcro2",5), rep("SeptAcro3",4), rep("SeptVeg1",3), rep("SeptVeg2",2), rep("AugVeg1",3), rep("AugVeg2",3), rep("AugVeg3",5), rep("AugAcro3",4), rep("SeptVeg3",4), rep("OctVeg3",3), rep("OctAcro3",3), rep("OctFem1",3), rep("OctFem2",3), rep("OctFem3",2), rep("AugVeg4",4), rep("SeptVeg4",4))

imgs <- c(
rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide1_V19T26-100_A1_SeptFem1.jpg",2),    rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide1_V19T26-100_B1_AugFem1.jpg",3), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide1_V19T26-100_C1_AugFem2.jpg",3), 
rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide1_V19T26-100_D1_AugFem2.jpg",3),
rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide2_V19T26-029_A1_SeptFem2.jpg",2), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide2_V19T26-029_B1_SeptFem3.jpg",2), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide2_V19T26-029_C1_AugAcro1_2.jpg",4), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide2_V19T26-029_D1_SeptAcro1.jpg",4), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide3_V19T26-078_A1_OctVeg1.jpg",3), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide3_V19T26-078_B1_OctVeg2.jpg",3), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide3_V19T26-078_C1_OctAcro1.jpg",2), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide3_V19T26-078_D1_OctAcro2.jpg",2), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide4_V19T26-079_A1_SeptAcro2.jpg",5), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide4_V19T26-079_B1_SeptAcro3.jpg",4), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide4_V19T26-079_C1_SeptVeg1_2.jpg",5), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide4_V19T26-079_D1_AugVeg1_2.jpg",6), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide5_V10F24-116_B1_AugVeg3.jpg",5), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide6_V10F24-117_A1_AugAcro3.jpg",4), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide6_V10F24-117_B1_SeptVeg3.jpg",4), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide6_V10F24-117_C1_OctVeg3.jpg",3), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide6_V10F24-117_D1_OctAcro3.jpg",3), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide7_V10F24-109_A1_OctFem1.jpg",3), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide7_V10F24-109_B1_OctFem2.jpg",3), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide7_V10F24-109_C1_OctFem3.jpg",2), rep("/Users/yuvarani.masarapu/Documents/23-02-2021_spruce_new/Spruce_HE_images/Slide7_V10F24-109_D1_AugVeg4_SeptVeg4.jpg",8))
  

infoTable <- data.frame(samples, 
                        spotfiles, 
                        imgs, 
                        section_id,
                        array_id,
                        slide_id,
                        timepoint,
                        stage,
                        bud,
                        stringsAsFactors = F)
```

## Loading data to seurat objects

### Load raw data *without filtering*

```{r}
se <- InputFromTable(infotable = infoTable, 
                                 platform = "Visium")

saveRDS(se, "/home/st-analysis_home/sami.saarenpaa/runs/Spruce/210222_spruce_raw_data.rds")
se <- LoadImages(se, time.resolve = F, verbose = T)
```

### Load raw data *with filtering*

This is a special step, instead of applying filtering on the unfiltered seurat object from above chunk, we re-do InputFromTable() with an extra filtering parameters defined. This makes sure that spots with low read count are recovered and not lost during the analysis.

Not sure why, but this (i.e the loss of spots from the bud sections) happened when the seurat object from above chunk was used directly for filtering low count spots and the for normalisation. Hence, backtracking was done to re-do this step as follow.

Read each row from data frame and :

1. split data frame into a list of rows
2. capture output when each row is converted into a seurat object, filtering step applied here; can be used later for statistics purpose.

```{r}
#re-order rows based on timepoint
infoTable <- infoTable[order(infoTable$timepoint), ]

#step 1; split data frame into a list of individual rows for better application
infoTable.list <- setNames(split(infoTable, seq(nrow(infoTable))), rownames(infoTable))

#step 2
capture <- capture.output(infoTable.counts.sep <- lapply(infoTable.list, function(x)  InputFromTable(x, 
                      platform="Visium",
                      transpose = F, 
                      min.spot.feature.count = 100,  min.spot.count = 200) ))

saveRDS(infoTable.counts.sep, file = "infotable_counts.rds")
```