---
title: "Marker genes analysis"
author: "Yuvarani Masarapu"
date: "4/9/2021"
output: html_document
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.path='figs/', fig.width=14, fig.height=10, warning=FALSE, message=FALSE, echo = FALSE)
```

```{r libraries-load, echo=FALSE}
suppressPackageStartupMessages({
  library(STutility)
  library(Seurat)
  library(harmony)
  library(gridExtra)
  library(pals)
  library(akima)
  library(readr)
  library(plotly)
  library(tibble)
  library(raster)
  library(dplyr)
  library(stringr)
  library(magrittr)
  library(ggplot2)
  library(imager)
  library(Matrix)
  library(cowplot)
  require(data.table)
  library(magick)
  library(grid)
  library(SeuratObject)
})
```

Venn diagram sets applied to find cluster specific markers and markers common to:

1. pith clusters:0,1,6,7
2. Vascular clusters:3,4,9,10,13
3. lateral organs' clusters: 2,11

OBS: Used FindAllMarkers() instead of FindMarkers() because in the latter, markers are forced to only one group so markers from closely related groups can be missed which the former tends to keep.

https://www.biostars.org/p/409790/

```{r}
#Read the csv file which contain markers from all clusters
library(readr)
AllMarkers <- read_csv("~/Documents/23-02-2021_spruce_new/AllMarkers.csv")

#Remove the first column which is the same as last column
AllMarkers <- AllMarkers[-1]

#save markers from each cluster as a separate list/vector
clus0 <- subset(AllMarkers, AllMarkers$cluster == 0)
saveRDS(clus0, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/clus0.rds")

clus1 <- subset(AllMarkers, AllMarkers$cluster == 1)
saveRDS(clus1, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/clus1.rds")

clus2 <- subset(AllMarkers, AllMarkers$cluster == 2)
saveRDS(clus2, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/clus2.rds")

clus3 <- subset(AllMarkers, AllMarkers$cluster == 3)
saveRDS(clus3, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/clus3.rds")

clus4 <- subset(AllMarkers, AllMarkers$cluster == 4)
saveRDS(clus4, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/clus4.rds")

clus5 <- subset(AllMarkers, AllMarkers$cluster == 5)
saveRDS(clus5, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/clus5.rds")

clus6 <- subset(AllMarkers, AllMarkers$cluster == 6)
saveRDS(clus6, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/clus6.rds")

clus7 <- subset(AllMarkers, AllMarkers$cluster == 7)
saveRDS(clus7, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/clus7.rds")

clus8 <- subset(AllMarkers, AllMarkers$cluster == 8)
saveRDS(clus8, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/clus8.rds")

clus9 <- subset(AllMarkers, AllMarkers$cluster == 9)
saveRDS(clus9, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/clus9.rds")

clus10 <- subset(AllMarkers, AllMarkers$cluster == 10)
saveRDS(clus10, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/clus10.rds")

clus11 <- subset(AllMarkers, AllMarkers$cluster == 11)
saveRDS(clus11, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/clus11.rds")

clus12 <- subset(AllMarkers, AllMarkers$cluster == 12)
saveRDS(clus12, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/clus12.rds")

clus13 <- subset(AllMarkers, AllMarkers$cluster == 13)
saveRDS(clus13, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/clus13.rds")

clus14 <- subset(AllMarkers, AllMarkers$cluster == 14)
saveRDS(clus14, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/clus14.rds")

clus15 <- subset(AllMarkers, AllMarkers$cluster == 15)
saveRDS(clus15, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/clus15.rds")


all.except.pith <- subset(AllMarkers, AllMarkers$cluster == c(2:5, 8:15))

all.except.vascular <- subset(AllMarkers, AllMarkers$cluster == c(0:2, 5:8, 11:12))

all.except.lateral <- subset(AllMarkers, AllMarkers$cluster == c(0:1, 3:10, 12:15))

common_pith <- intersect(intersect(intersect(clus0$gene, clus1$gene), clus6$gene), clus7$gene)
saveRDS(common_pith, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/common_pith.rds")

common_vascular <- intersect(intersect(intersect(intersect(clus3$gene, clus4$gene), clus9$gene), clus10$gene), clus13$gene)
saveRDS(common_vascular, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/common_vascular.rds")

common_lateral <- intersect(clus2$gene, clus11$gene)
saveRDS(common_lateral, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/common_lateral.rds")

#find unique pith markers in a vector
j=1
count=0
vec.uniq.pith <- vector(mode = "character")
for (i in 1:length(common_pith)){
    if (common_pith[i] %in% all.except.pith$gene)
    {
      count <- count+1
    }
    else{
      vec.uniq.pith[j] <- common_pith[i]
      j = j+1
    }
}
saveRDS(vec.uniq.pith, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/uniq_pith_genes.rds")

#find unique vascular markers in a vector
j=1
count=0
vec.uniq.vascular <- vector(mode = "character")
for (i in 1:length(common_vascular)){
    if (common_vascular[i] %in% all.except.vascular$gene)
    {
      count <- count+1
    }
    else{
      vec.uniq.vascular[j] <- common_vascular[i]
      j = j+1
    }
}
saveRDS(vec.uniq.vascular, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/uniq_vascular_genes.rds")

#find unique lateral markers in a vector
j=1
count=0
vec.uniq.lateral <- vector(mode = "character")
for (i in 1:length(common_lateral)){
    if (common_lateral[i] %in% all.except.lateral$gene)
    {
      count <- count+1
    }
    else{
      vec.uniq.lateral[j] <- common_lateral[i]
      j = j+1
    }
}
saveRDS(vec.uniq.lateral, file = "~/Documents/23-02-2021_spruce_new/markers_clusters/uniq_lateral_genes.rds")


#save all the information in an excel sheet
list_of_datasets <-
  list("cluster0" = clus0, "cluster1" = clus1, "cluster2" = clus2, "cluster3" = clus3, "cluster4" = clus4, "cluster5" = clus5, "cluster6" = clus6, "cluster7" = clus7, "cluster8" = clus8, "cluster9" = clus9, "cluster10" = clus10, "cluster11" = clus11, "cluster12" = clus12, "cluster13" = clus13, "cluster14" = clus14, "cluster15" = clus15, 
"pith_common" = data.frame("genes" = common_pith), "vascular_common" = data.frame("genes" = common_vascular), "lateral_common" = data.frame("genes" = common_lateral), "pith_unique" = data.frame("genes" = vec.uniq.pith), "vascular_unique" = data.frame("genes" = vec.uniq.vascular), "lateral_unique" = data.frame("genes" = vec.uniq.lateral))

write_xlsx(list_of_datasets,"~/Documents/23-02-2021_spruce_new/markers_clusters/markers_in_clusters.xlsx")
```


Find bud specific markers for cluster 15 which was troublesome in the biological analysis

*OBS!* Set "min.cells.group = 1" if you get an error 
  Error in ValidateCellGroups(object = object, cells.1 =         cells.1, cells.2 = cells.2,  : 
  Cell group 1 has fewer than 3 cells
  
This makes sure to set minimum number of spots in the cluster to 1, otherwise the function FindMarkers() by default takes minimum number of spots to be 3 and throws an error if the samples contain less than 3 spots for the particular cluster we want to find markers for.

Discussed here: https://github.com/satijalab/seurat/issues/1001

```{r}
spruce <- readRDS(file = "~/Documents/23-02-2021_spruce_new/spruce_merged_PCA_harmony_Umap_FindNeigh_findClusters.rds")

octAcro.15 <- FindMarkers(spruce, ident.1 = c("V19T26-078-C1-2", "V19T26-078-D1-1", "V10F24-117-D1-1"), group.by = c("section_id"), subset.ident = "15", test.use = "bimod")

#octAcro.1 <- FindMarkers(spruce, ident.1 = "V19T26-078-C1-2", group.by = "section_id")

octFem.15 <- FindMarkers(spruce, ident.1 = c("V10F24-109-A1-1", "V10F24-109-A1-2", "V10F24-109-A1-3", "V10F24-109-B1-1", "V10F24-109-B1-2", "V10F24-109-B1-3", "V10F24-109-C1-1", "V10F24-109-C1-2"), group.by = c("section_id"), subset.ident = "15", test.use = "bimod")

octVeg.15 <- FindMarkers(spruce, ident.1 = c("V19T26-078-A1-2", "V19T26-078-B1-1"), group.by = c("section_id"), subset.ident = "15", test.use = "bimod", min.cells.group = 1)

septAcro.15 <- FindMarkers(spruce, ident.1 = c("V19T26-029-D1-3", "V19T26-079-A1-3", "V19T26-079-A1-4", "V19T26-079-B1-3"), group.by = c("section_id"), subset.ident = "15", test.use = "bimod")

septFem.15 <- FindMarkers(spruce, ident.1 = c("V19T26-029-B1-2"), group.by = c("section_id"), subset.ident = "15", test.use = "bimod", min.cells.group = 1)


#save all in excel sheet
octAcro.15$gene <- row.names(octAcro.15)
octVeg.15$gene <- row.names(octVeg.15)
octFem.15$gene <- row.names(octFem.15)
septAcro.15$gene <- row.names(septAcro.15)
septFem.15$gene<- row.names(septFem.15)

list_of_datasets <-
  list("OctAcro" = octAcro.15, "OctVeg" = octVeg.15, "OctFem" = octFem.15, "SeptAcro" = septAcro.15, "SeptFem" = septFem.15)

write_xlsx(list_of_datasets,"~/Documents/23-02-2021_spruce_new/markers_clusters/cluster15_tissueSpecific_markers.xlsx")
```